
\chapter{Star Catalogues}
\label{ch:Catalogues}

This chapter provides technical descriptions on how Stellarium records its star catalogues, and
the related file formats.

\section{Stellarium's Sky Model}
\label{sec:Catalogues:SkyModel}

\subsection{Zones}
\label{sec:Catalogues:SkyModel:Zones}

The celestial sphere is split into zones, which correspond to the
triangular faces of a geodesic sphere. The number of zones (faces)
depends on the level of sub-division of this sphere plus one global zone for 
high proper motion stars that can cross zones in the past/future or stars. The lowest level,
0, is an icosahedron (20 faces + 1 global zone), subsequent levels $L$ of sub-division
give the number of zones $n$ as:

\begin{equation}
n=20 \cdot 4^L + 1
\end{equation}

Stellarium uses levels 0 to 8 in the existing star catalogues. So the number of zones are 21 for level 0 to 1,310,721 for level 8.
Stars within the global zone are always checked if they are visible or not without any zone restriction, since stars can move 
out of their original zones due to proper motion and not drawn in the sky. Only level 0 to level 2 star catalogs have their global 
zone populated (i.e. level 3-8 global zones are empty).

For level 0, any stars who will move out of their original zones in $\pm 100,000$ years are moved to the global zone. For level 1 and 2, stars are moved 
to the global zone only if they move out of their original zones and the bordering zones \textbf{or} brighter than 0.3 mag from their 
present apparent magnitude in $\pm 100,000$ years.

\section{Star Catalogue File Format}%\label{star-catalogue-file-format}
\label{sec:Catalogues:stars}

With version 25.1\newFeature{25.1}, Stellarium is adopting the currently most modern class of astrometrical data: Gaia DR3. 
This astrometry satellite, operated by ESA, delivers unprecedented accuracy for millions of stars. 
However, just those stars most important to human casual observers, naked-eye stars, 
are too bright for the sensitive instruments. Therefore, for the brightest stars, 
data from the preceding Hipparcos mission are still used. Most indexing is still based on Hipparcos numbers. 
In previous versions, dimmer stars were stored without number. Now, you can use the long index from Gaia DR3 
in descriptions of constellations, or search for the number directly. 

Information about the previous format is easiest found in the v24.4 User Guide.
 

\subsection{General Description}%\label{general-description}
\label{sec:Catalogues:stars:general}

Stellarium's star catalogue data is kept in the \file{stars/hip\_gaia3}\newFeature{25.1}
sub-directory of the Installation Directory and/or User Directory (see
section~\ref{sec:Directories}).\footnote{The data has significantly changed in version 25.1. The previous structure can be found in older guides.}

The main catalogue data is split into several files:

\begin{description}
\item[\file{stars\_0\_0v0\_18.cat}]
\item[\file{stars\_1\_0v0\_15.cat}]
\item[\file{stars\_2\_0v0\_16.cat}]
\item[\file{stars\_3\_0v0\_10.cat}]
\item[\file{stars\_4\_1v0\_6.cat}]
\item[\file{stars\_5\_1v0\_6.cat}]
\item[\file{stars\_6\_1v0\_4.cat}]
\item[\file{stars\_7\_1v0\_4.cat}]
\item[\file{stars\_8\_2v0\_3.cat}]
\end{description}

There also exist some control and reference files:

\begin{description}
\item[\file{object\_types\_v0\_1.cat}]
\item[\file{stars\_hip\_sp\_0v0\_6.cat}]
\item[\file{gcvs.cat}]
\item[\file{wds\_hip\_part.dat}]
\item[\file{cross-id.cat}]
\item[\file{starsConfig.json}]
\item[\file{name.fab}]
\item[\file{binary\_orbitparam.dat}]
\end{description}

When Stellarium starts, it reads the file \file{stars/hip\_gaia3/starsConfig.json}%
\footnote{Or \file{stars/hip\_gaia3/defaultStarsConfig.json} file in Installation Directory}, from which it determines the 
names of the other files which it then loads. The file \file{stars\_hip\_sp\_0v0\_6.cat} 
contain spectral type of HIP stars. The file \file{object\_types\_v0\_1.cat} 
contain object type classification name for the main star catalogs. 
The file \file{gcvs.cat} contains data about variables stars
(see section~\ref{sec:StarCatalogues:VariableStars}) and file \file{wds\_hip\_part.dat} 
contains data about double stars (see section~\ref{sec:StarCatalogues:DoubleStars}) 
which again contains references into the main catalogue files.
The file \file{cross-id.cat} (see section~\ref{sec:StarCatalogues:CrossIdentificationData}) contains cross-identification data between HIP, Gaia DR3, SAO, HD and HR designations.
The file \file{binary\_orbitparam.cat} (see section~\ref{sec:StarCatalogues:BinaryOrbitalParameters}) contains binary star system orbital parameters.

A given catalogue file models stars for one and only one level (i.e. for
a fixed number of zones), which is recorded in the header of the file.
Individual star records do not contain full positional coordinates,
instead they contain coordinates relative to the central position of the
zone they occupy. Thus, when parsing star catalogues, it is necessary to
know about the zone model to be able to extract positional data.

\noindent
\begin{tabular}{ccrcrl}
\toprule
\emph{File} & \emph{Data     } & \emph{Data            } & \emph{Geodesic      } & \emph{\#Records} & \emph{Notes}\\
            & \emph{     Type} & \emph{     Record Size} & \emph{         Level} &                  &             \\
\midrule
\file{stars\_0\_0v0\_18.cat} & 0 & 48 bytes & 0 &        5,045 & Hipparcos + Gaia\\
\file{stars\_1\_0v0\_15.cat} & 0 & 48 bytes & 1 &       21,611 & Hipparcos + Gaia\\
\file{stars\_2\_0v0\_16.cat} & 0 & 48 bytes & 2 &      141,761 & Hipparcos + Gaia\\
\file{stars\_3\_0v0\_10.cat} & 0 & 48 bytes & 3 &      418,129 & Gaia\\
\file{stars\_4\_1v0\_6.cat}  & 1 & 32 bytes & 4 &    1,741,849 & Gaia\\
\file{stars\_5\_1v0\_6.cat}  & 1 & 32 bytes & 5 &    8,051,933 & Gaia\\
\file{stars\_6\_1v0\_4.cat}  & 1 & 32 bytes & 6 &   29,879,734 & Gaia\\
\file{stars\_7\_1v0\_4.cat}  & 1 & 32 bytes & 7 &   57,534,724 & Gaia\\
\file{stars\_8\_2v0\_3.cat}  & 2 & 16 bytes & 8 &  122,908,184 & Gaia\\
\bottomrule
\end{tabular}

For a given catalogue file, there may be one of three formats for the
actual star data. The variation comes from the source of the data -- the
larger catalogues of fainter stars providing less data per star than the
brighter star catalogues. See tables \emph{Stellarium's star catalogue}
and for details.

Stellarium's star catalogues are based on Hipparcos~\citep{1997ESASP1200.....E, 2012AstL...38..331A} and Gaia~\citep{2016A&A...595A...1G} catalogues.

\subsection{File Sections}%\label{file-sections}
\label{sec:Catalogues:stars:sections}

The catalogue files are split into three main sections as described in
table~\ref{tab:Catalogues:stars:sections}.

\begin{table}[htb]
\begin{tabularx}{\textwidth}{l l X}\toprule
\emph{Section} & \emph{Offset} & \emph{Description}\\\midrule
File Header Record &  0        & Contains magic number, geodesic subdivision level, magnitude and epoch information\\%\midrule
Zone Records       & 28        & A list of how many records there are for each zone.
                                 The length of the zones section depends on the level value from the header\\%\midrule
Star Data Records  & $28+4n$   & This section of the file contains fixed-size star records, as described below. 
                                 Records do not contain zone information, which must be inferred by counting 
                                 how many records have been read so far and switching zones when enough have been read 
                                 to fill the number of stars for the zone, as specified in the zones section above. 
                                 The value of $n$ used in the offset description is the number of zones, as described above.\\\bottomrule
\end{tabularx}
\caption{File sections}
\label{tab:Catalogues:stars:sections}
\end{table}

\subsection{Record Types}%\label{record-types}
\label{sec:Catalogues:stars:record:types}

\subsubsection{File Header Record}%\label{file-header-record}
\label{sec:Catalogues:stars:record:header}

The File Header Record describes file-wide settings. It also contains a
magic number which servers as a file type identifier. See table~\ref{tab:Catalogues:stars:record:header}.

\begin{table}[htbp]
\begin{tabularx}{\textwidth}{llllX}\toprule
\emph{Name} & \emph{Offset} & \emph{Type} & \emph{Size} & \emph{Description}\\\midrule
Magic             &  0 & int   & 4 & magic number which identifies the file as a star catalogue. \texttt{0x835f040a}\\%\midrule
Data Type         &  4 & int   & 4 & type of the file, which defines the size and structure 
                                   of the Star Data record for the file.\\%\midrule
Major Version     &  8 & int   & 4 & file format major version number\\%\midrule
Minor Version     & 12 & int   & 4 & file format minor version number\\%\midrule
Level             & 16 & int   & 4 & sets the level of sub-division of the geodesic sphere used to create the zones. 
                                   0 means an icosahedron (20 triangular faces + 1 global zone), 
                                   subsequent levels of sub-division lead to numbers of zones 
                                   as described in section \ref{sec:Catalogues:stars:record:zone}\\%\midrule
Magnitude Minimum & 20 & int   & 4 &  low bound of the magnitude of the catalog in milli-mag unit.  \\%\midrule
Catalog Epoch     & 24 & float & 4 &  Epoch of the catalog in Julian days.\\%\midrule
\end{tabularx}
\caption{Header Record}
\label{tab:Catalogues:stars:record:header}
\end{table}

\subsubsection{Zone Records}%\label{zone-records}
\label{sec:Catalogues:stars:record:zone}

The \emph{Zone Records} section of the file lists the number of star
records there are per zone. The number of zones is determined from the
level value in the File Header Record, as described in
section~\ref{sec:Catalogues:SkyModel:Zones}. The Zones section is
simply a list of integer values which describe the number of stars for
each zone. The total length of the Zones section depends on the number
of zones. See table~\ref{tab:Catalogues:stars:record:zones}.


\subsubsection{Star Data Records}%\label{star-data-records}
\label{sec:Catalogues:stars:record:data}

After the Zones section, the actual star data starts. The star data
records themselves do not contain the zone in which the star belongs.
Instead, the zone is inferred from the position of the record in the
file. For example, if the Zone Records section of the file says that the
first 100 records are for zone 0, the next 80 for zone 1 and so on, it
is possible to infer the zone for a given record by counting how many
records have been read so far.

The actual record structure depends on the value of the Data Type, as
found in the File Header Record.
See tables 
%\emph{Star Data Record Type 0}, \emph{Star Data Record Type 1} and \emph{Star Data Record Type 2} 
\ref{tab:StarDataRecord0}, \ref{tab:StarDataRecord1} and \ref{tab:StarDataRecord2}
for record structure details.

Details about vector algebra related to astrometric calculations can be 
found in \citep{2012asas.book.....V}. For the catalog with Star Data Record 
Type 0, in practice to convert $\mathtt{x0}, \mathtt{x1}, \mathtt{x2}$ to 
right ascension $\alpha$ and declination $\delta$, one can use the following 
formulas:

\begin{equation}
\begin{aligned}
\alpha &= \arctan2(\mathtt{x1}, \mathtt{x0})\\
\delta &= \arcsin(\mathtt{x2})
\end{aligned}
\end{equation}

and to convert proper motion vector $(\mathtt{dx0}, \mathtt{dx1}, \mathtt{dx2})$ to 
proper motion in right ascension $\mu_{\alpha^*}$ and declination $\mu_{\delta}$, 
one can use the following formulas:

\begin{equation}
\begin{aligned}
	xy &= \sqrt{\mathtt{x0}^2 + \mathtt{x1}^2}; \\
	\vec{p} &= \left(-\frac{\mathtt{x1}}{xy}, \frac{\mathtt{x0}}{xy}, 0.\right); \\
	\vec{q} &= \left(-\frac{\mathtt{x0} \cdot \mathtt{x2}}{xy}, -\frac{\mathtt{x1} \cdot \mathtt{x2}}{xy}, xy\right);\\
	\mu_{\alpha^*} &= \vec{p} \cdot \left(\mathtt{dx0}, \mathtt{dx1}, \mathtt{dx2}\right); \\
	\mu_{\delta} &= \vec{q} \cdot \left(\mathtt{dx0}, \mathtt{dx1}, \mathtt{dx2}\right);
\end{aligned}
\end{equation}

where $\mu_{\alpha^*}\equiv\mu_{\alpha}\cos{(\delta)}$.

\begin{table}[htbp]
\begin{tabularx}{\textwidth}{llllX}\toprule
\emph{Name} & \emph{Offset} & \emph{Type} & \emph{Size} &\emph{Description}\\\midrule
num stars in zone 0     & 0      & int & 4 &  number of records in this file which are in zone 0\\
num stars in zone 1     & 4      & int & 4 &  number of records is this file which are in zone 1\\
\ldots                  &        &     &   &\\
num stars in zone $n-1$ & $4n-4$ & int & 4 &  number of records is this file which are in zone $n$\\\bottomrule
num stars in zone $n$   & $4n$   & int & 4 &  number of records is this file which are in the global zone\\\bottomrule
\end{tabularx}
\caption{Zones section}
\label{tab:Catalogues:stars:record:zones}
\end{table}


\begin{table}[htbp]
\begin{tabular}{llllp{75mm}}\toprule
\emph{Name} & \emph{Offset} & \emph{Type} & \emph{Size} &\emph{Description}\\\midrule
gaia\_id       &  0 & int64\_t           & 8 & Gaia source id\\%\midrule
x0             &  8 & int                & 4 & Cartesian x position of the star on an unit sphere at the catalog epoch scaled by $2\times10^9$\\
x1             & 12 & int                & 4 & Cartesian y position of the star on an unit sphere at the catalog epoch scaled by $2\times10^9$\\
x2             & 16 & int                & 4 & Cartesian z position of the star on an unit sphere at the catalog epoch scaled by $2\times10^9$\\%\midrule
dx0            & 20 & int                & 4 & First component of the star's proper motion vector at the catalog epoch scaled by $1\times10^3$\\
dx1            & 24 & int                & 4 & Second component of the star's proper motion vector at the catalog epoch scaled by $1\times10^3$\\
dx2            & 28 & int                & 4 & Third component of the star's proper motion vector at the catalog epoch scaled by $1\times10^3$\\%\midrule
b\_v           & 32 & short              & 2 & $B-V$ colour in milli-mag unit.\\
mag            & 34 & short              & 2 & $V$-band magnitude in milli-mag unit.\\
plx            & 36 & unsigned short     & 2 & parallax of the star in 0.02 mas unit.\\
plx\_err       & 38 & unsigned short     & 2 & parallax error of the star in 0.01 mas unit.\\%\midrule
rv             & 40 & short              & 2 & radial velocity of the star in 0.1 km/s unit.\\%\midrule
sp\_int        & 42 & unsigned short     & 2 & index into an array of spectral type descriptions 
                                               which is taken from the file \file{stars\_hip\_sp.cat}, the index corresponds to the
                                               line number in the file - 1\\%\midrule
obtype	       & 44 & unsigned char      & 1 & object type classification index. The index corresponds to the line number in the file \file{object\_types\_v*.cat} - 1.\\%\midrule
hip            & 45 & int                & 3 & Hipparcos catalogue number with component ID (e.g. A, B, C...)\\\bottomrule
\end{tabular}
\caption{Star Data Record Type 0}
\label{tab:StarDataRecord0}
\end{table}

\begin{table}[htbp]
\begin{tabularx}{\textwidth}{llllX}\toprule
\emph{Name} & \emph{Offset}& \emph{Type} & \emph{Size} &\emph{Description}\\\midrule
gaia\_id       &  0 & int64\_t           & 8 & Gaia source id\\%\midrule
x0             &  8 & int                & 4 & Right ascension $\alpha$ of the star in mas unit.\\
x1             & 12 & int                & 4 & Declination $\delta$ of the star in mas unit.\\%\midrule
dx0            & 16 & int                & 4 & Proper motion $\mu_{\alpha}$ of the star in $\mu$as/year\\
dx1            & 20 & int                & 4 & Proper motion $\mu_{\delta}$ of the star in $\mu$as/year\\%\midrule
b\_v           & 24 & short              & 2 & $B-V$ colour in milli-mag unit.\\
mag            & 26 & short              & 2 & $V$-band magnitude in milli-mag unit.\\
plx            & 28 & unsigned short     & 2 & parallax of the star in 0.01 mas unit.\\
plx\_err       & 30 & unsigned short     & 2 & parallax error of the star in 0.01 mas unit.\\\bottomrule
\end{tabularx}
\caption{Star Data Record Type 1}
\label{tab:StarDataRecord1}
\end{table}


\begin{table}[htbp]
\begin{tabularx}{\textwidth}{llllX}\toprule
\emph{Name} & \emph{Offset} & \emph{Type} & \emph{Size} & \emph{Description}\\\midrule
gaia\_id       &  0 & int64\_t           & 8 & Gaia source id\\
x0             &  8 & int                & 3 & Right ascension $\alpha$ of the star in 0.1 arcsecond unit.\\
x1             & 11 & int                & 3 & Declination $\delta$ of the star in 0.1 arcsecond unit.\\
b\_v           & 14 & unsigned short     & 1 & $B-V$ colour in 25 milli-mag unit offset by $+1$ mag.\\
mag            & 15 & unsigned short     & 1 & $V$-band magnitude in 20 milli-mag unit offset by $-16$ mag.\\\bottomrule
\end{tabularx}
\caption{Star Data Record Type 2}
\label{tab:StarDataRecord2}
\end{table}


\clearpage

\section{Variable Stars}
\label{sec:StarCatalogues:VariableStars}

Since version 0.12.2 Stellarium supports the subset of variable stars
from GCVS\footnote{\url{http://www.sai.msu.su/gcvs/gcvs/}} which have HIP
identifiers (i.e., stars from the Hipparcos catalog). Stellarium's Catalog
of Variable Stars is based on the \citetp{samus}.
%\textbf{General Catalogue of Variable Stars} 

\subsection{Variable Star Catalog File Format}
\label{sec:StarCatalogues:VariableStars:format}

\subsubsection{General Description}%\label{general-description}
\label{sec:StarCatalogues:VariableStars:general}

Stellarium's Variable Stars catalog data is kept
in the \file{stars/hip\_gaia3} sub-directory of the Installation Directory and/or
User Directory (see section~\ref{sec:Directories}).


\subsubsection{File Format}%\label{file-format}
\label{sec:StarCatalogues:VariableStars:file}

The \file{gcvs.cat} file contains data about 49800+ Variable Stars 
(see section~\ref{sec:Phenomena:variableStars}) which have HIP or Gaia DR3 identifiers and
are stored in a plain text format with tab delimiter (Table~\ref{tab:sec:StarCatalogues:VariableStars}).

\begin{table}[tb]
\begin{tabularx}{\textwidth}{l|l|X}\toprule
\emph{Name} & \emph{Type} & \emph{Description}\\\midrule
HIP/Gaia & int    & This is the HIP or Gaia DR3 identifier for the star. Used for reference to the main star catalogue.\\
GCVS     & string & This is the GCVS designation for the variable star.\\
Type     & string & Type of variability.\\
Max      & float  & Magnitude at maximum brightness.\\
MFlag    & int    & Magnitude flag code (0=No flag; 1="(" if max. magnitude is an amplitude; 
                    2="\textless{}" if max. magnitude is a bright limit; 
                    3="\textgreater{}" is max. magnitude is a faint limit).\\
Min I    & float  & First minimum magnitude or amplitude.\\
Min II   & float  & Second minimum magnitude.\\
V        & char   & The photometric system for magnitudes.\\
Epoch    & float  & Epoch for maximum or minimum light (in Julian days).\\
Period   & float  & Period for the variable star (in days).\\
M-m      & float  & Rising time or duration of the eclipse (in percent).\\
Spectrum & string & Spectral class of the variable star.\\\bottomrule
\end{tabularx}
\caption{Variable Star Catalog File Format}
\label{tab:sec:StarCatalogues:VariableStars}
\end{table}

\subsection{GCVS Variability Types}\label{gcvs-variability-types}

An improved system of variability classification is used in the fourth
edition of the GCVS, based on recent developments in classification
principles and taking into account the suggestions of a number of
specialists. Variability types are grouped according to the major
astrophysical reasons for variability.

All of these classes include objects of a dissimilar nature that belong
to different types of light variability. On the other hand, an object
may be variable because of almost all of the possible reasons or because
of any combination of them. If a variable belongs to several types of
variability, the types are joined in the data field by a ``+'' sign, e.g.,
E+UG, UV+BY.

Despite considerable success in understanding stellar variability
processes, the classification adopted in the Catalogue is far from
perfect. This is especially the case for explosive, symbiotic and
novalike variables; X-ray sources; and peculiar objects.

The new variability types (ZZO, AM, R, BE, BLBOO, EP, SRS, LPB) have been 
added in the Name-Lists 67-77 and in the GCVS vol.V.

You can read about latest different types of variability at SAI GCVS page\footnote{
\url{http://www.sai.msu.su/gcvs/gcvs/vartype.htm}}.

\subsubsection{Eruptive Variable Stars}\label{eruptive-variable-stars}

Eruptive variables are stars varying in brightness because of violent
processes and flares occurring in their chromospheres and coronae. The
light changes are usually accompanied by shell events or mass outflow in
the form of stellar winds of variable intensity and/or by interaction
with the surrounding interstellar medium. This class includes the
following types:

\begin{description}
\item[BE]       It becomes more and more clear that, although the majority
                of Be stars are photometrically variable, not all of them could be
                properly called GCAS variables. Quite a number of them show small-scale
                variations not necessarily related to shell events; in some cases the
                variations are quasi-periodic. By now we are not able to present an
                elaborated system of classification for Be variables, but we adopt a
                decision that in the cases when a Be variable cannot be readily
                described as a GCAS star we give simply BE for the type of variability.
\item[FU]		Orion variables of the FU Orionis type. Characterized by
				gradual increases in brightness by about 6 mag in several months,
				followed by either almost complete constancy at maximum that is
				sustained for long periods of time or slow decline by 1-2 mag. Spectral
				types at maximum are in the range Ae(alpha) - Gpe(alpha). After an
				outburst, a gradual development of an emission spectrum is observed and
				the spectral type becomes later. These variables probably mark one of
				the evolutionary stages of T Tauri-type Orion variables (INT), as
				evidenced by an outburst of one member, V1057 Cyg, but its decline (2.5
				mag in 11 years) commenced immediately after maximum brightness was
				attained. All presently known FU Ori variables are coupled with
				reflecting cometary nebulae.
\item[GCAS]		Eruptive irregular variables of the Gamma Cas type.
				These are rapidly rotating B III-IVe stars with mass outflow from their
				equatorial zones. The formation of equatorial rings or disks is often
				accompanied by temporary fading. Light amplitudes may reach 1.5 mag in V.
\item[I]		Poorly studied irregular variables with unknown features of
				light variations and spectral types. This is a very inhomogeneous group
				of objects.
	\begin{description}
\item[IA] 		Poorly studied irregular variables of early (O-A) spectral
				type.
\item[IB]		Poorly studied irregular variables of intermediate (F-G)
				to late (K-M) spectral type.
\item[IN]		Orion variables. Irregular, eruptive variables connected
				with bright or dark diffuse nebulae or observed in the regions of these
				nebulae. Some of them may show cyclic light variations caused by axial
				rotation. In the Spectrum-Luminosity diagram, they are found in the area
				of the main sequence and subgiants. They are probably young objects
				that, during the course of further evolution, will become light-constant
				stars on the zero-age main sequence (ZAMS). The range of brightness
				variations may reach several magnitudes. In the case of rapid light
				variations having been observed (up to 1 mag in 1-10 days), the letter
				``S'' is added to the symbol for the type (INS). This type may be
				divided into the following subtypes:
		\begin{description}
\item[INA]		Orion variables of early spectral types (B-A or Ae). They
				are often characterized by occasional abrupt Algol-like fadings (T Ori);
\item[INB]		Orion variables of intermediate and late spectral types,
				F-M or Fe-Me (BH Cep, AH Ori). F-type stars may show Algol-like fadings
				similar to those of many INA stars; K-M stars may produce flares along
				with irregular light variations;
\item[INT,IT] 	Orion variables of the T Tauri type. Stars are assigned
				to this type on the basis of the following (purely spectroscopic)
				criteria: spectral types are in the range Fe-Me. The spectra of most
				typical stars resemble the spectrum of the solar chromosphere. The
				feature specific to the type is the presence of the fluorescent emission
				lines Fe II 4046, 4132 A (anomalously intense in the spectra of these
				stars), emission lines {[}Si II{]} and {[}O I{]}, as well as the
				absorption line Li I 6707 A. These variables are usually observed only
				in diffuse nebulae. If it is not apparent that the star is associated
				with a nebula, the letter ``N'' in the symbol for the type may be
				omitted, e.g., IT (RW AUR);
\item[IN(YY)] 	Some Orion variables (YY Ori) show the presence of
				absorption components on the redward sides of emission lines, indicating
				the infall of matter toward the stars' surfaces. In such cases, the
				symbol for the type may be accompanied by the symbol
				``YY''.
		\end{description}
\item[IS]		Rapid irregular variables having no apparent connection
				with diffuse nebulae and showing light changes of about 0.5 - 1.0 mag
				within several hours or days. There is no strict boundary between rapid
				irregular and Orion variables. If a rapid irregular star is observed in
				the region of a diffuse nebula, it is considered an Orion variable and
				designated by the symbol INS. To attribute a variable to the IS type, it
				is necessary to take much care to be certain that its light changes are
				really not periodic. Quite a number of the stars assigned to this type
				in the third edition of the GCVS turned out to be eclipsing binary
				systems, RR Lyrae variables, and even extragalactic BL Lac
				objects.
		\begin{description}
\item[ISA]		Rapid irregular variables of the early spectral types, B-A
				or Ae;
\item[ISB]		Rapid irregular variables of the intermediate and late
				spectral types, F-M and Fe-Me.
		\end{description}
	\end{description}
\item[RCB]		Variables of the R Coronae Borealis type. These are
				hydrogen-poor, carbon- and helium-rich, high-luminosity stars belonging
				to the spectral types Bpe-R, which are simultaneously eruptive and
				pulsating variables. They show slow nonperiodic fadings by 1-9 mag in V
				lasting from a month or more to several hundred days. These changes are
				superposed on cyclic pulsations with amplitudes up to several tenths of
				a magnitude and periods in the range 30-100 days.
\item[RS]		Eruptive variables of the RS Canum Venaticorum type. This
				type is ascribed to close binary systems with spectra showing Ca II H
				and K in emission, their components having enhanced chromospheric
				activity that causes quasi-periodic light variability. The period of
				variation is close to the orbital one, and the variability amplitude is
				usually as great as 0.2 mag in V (UX Ari). They are X-ray sources and
				rotating variables. RS CVn itself is also an eclipsing system (see
				below).
\item[SDOR]		Variables of the S Doradus type. These are eruptive,
				high-luminosity Bpec-Fpec stars showing irregular (sometimes cyclic)
				light changes with amplitudes in the range 1-7 mag in V. They belong to
				the brightest blue stars of their parent galaxies. As a rule, these
				stars are connected with diffuse nebulae and surrounded by expanding
				envelopes (P Cyg, Eta Car).
\item[UV]		Eruptive variables of the UV Ceti type, these are K Ve-M
				Ve stars sometimes displaying flare activity with amplitudes from
				several tenths of a magnitude up to 6 mag in V. The amplitude is
				considerably greater in the ultraviolet spectral region. Maximum light
				is attained in several seconds or dozens of seconds after the beginning
				of a flare; the star returns to its normal brightness in several minutes
				or dozens of minutes.
\item[UVN]		Flaring Orion variables of spectral types Ke-Me. These
				are phenomenologically almost identical to UV Cet variables observed in
				the solar neighborhood. In addition to being related to nebulae, they
				are normally characterized by being of earlier spectral type and greater
				luminosity, with slower development of flares (V389 Ori). They are
				possibly a specific subgroup of INB variables with irregular variations
				superimposed by flares.
\item[WR]		Eruptive Wolf-Rayet variables. Stars with broad emission
				features of He I and He II as well as C II-C IV, O II-O IV, and N III-N
				V. They display irregular light changes with amplitudes up to 0.1 mag in
				V, which are probably caused by physical processes, in particular, by
				nonstable mass outflow from their atmospheres.
\end{description}

\subsubsection{Pulsating Variable Stars}\label{pulsating-variable-stars}

Pulsating variables are stars showing periodic expansion and contraction
of their surface layers. The pulsations may be radial or nonradial. A
radially pulsating star remains spherical in shape, while in the case of
nonradial pulsations the star's shape periodically deviates from a
sphere, and even neighboring zones of its surface may have opposite
pulsation phases.

Depending on the period value, on the mass and evolutionary status of
the star, and on the scale of pulsational phenomena, the following types
of pulsating variables may be distinguished:

\begin{description}
\item[ACYG]		Variables of the Alpha Cygni type, which are nonradially
				pulsating supergiants of Bep-AepIa spectral types. The light changes
				with amplitudes of the order of 0.1 mag often seem irregular, being
				caused by the superposition of many oscillations with close periods.
				Cycles from several days to several weeks are observed.
\item[BCEP]		Variables of the Beta Cephei type (Beta Cep, Beta CMa),
				which are pulsating O8-B6 I-V stars with periods of light and
				radial-velocity variations in the range of 0.1 - 0.6 days and light
				amplitudes from 0.01 to 0.3 mag in V. The light curves are similar in
				shape to average radial-velocity curves but lag in phase by a quarter of
				the period, so that maximum brightness corresponds to maximum
				contraction, i.e., to minimum stellar radius. The majority of these
				stars probably show radial pulsations, but some (V469 Per) display
				nonradial pulsations; multiperiodicity is characteristic of many of
				these stars.
	\begin{description}
\item[BCEPS]	A short-period group of Beta Cep variables. The
				spectral types are B2-B3 IV-V; periods and light amplitudes are in the
				ranges 0.02 - 0.04 days and 0.015 - 0.025 days, respectively, i.e., an
				order of magnitude smaller than the normally observed ones.
	\end{description}
\item[BLBOO]    The so-called ``anomalous Cepheids'', i.e. stars with
                periods characteristic of comparatively long-period RRAB variables, but
                considerably brighter by luminosity (BL Boo = NGC 5466 V19).
\item[CEP]		Cepheids. Radially pulsating, high luminosity (classes
				Ib-II) variables with periods in the range of 1-135 days and amplitudes
				from several hundredths to 2 mag in V (in the B band, the amplitudes are
				greater). Spectral type at maximum light is F; at minimum, the types are
				G-K. The longer the period of light variation, the later is the spectral
				type. The maximum of the surface-layer expansion velocity almost
				coinciding with maximum light.
	\begin{description}
\item[CEP(B)]	Cepheids (TU Cas, V 367 Sct) displaying the presence
				of two or more simultaneously operating pulsation modes (usually the
				fundamental tone with the period P0 and the first overtone P1). The
				periods P0 are in the range from 2 to 7 days, with the ratio P1/P0
				approx. 0.71.
	\end{description}
\item[CW]		Variables of the W Virginis type. These are pulsating
				variables of the galactic spherical component (old disk) population with
				periods of approximately 0.8 to 35 days and amplitudes from 0.3 to 1.2
				mag in V. They obey a period-luminosity relation different from that for
				Delta Cep variables (see DCEP). For an equal period value, the W Vir
				variables are fainter than the Delta Cep stars by 0.7 - 2 mag. The light
				curves of W Vir variables for some period intervals differ from those of
				Delta Cep variables for corresponding periods either by amplitudes or by
				the presence of humps on their descending branches, sometimes turning
				into broad flat maxima. W Vir variables are present in globular clusters
				and at high galactic latitudes. They may be separated into the following
				subtypes:
	\begin{description}
\item[CWA]		W Vir variables with periods longer than 8 days (W Vir);
\item[CWB]		W Vir variables with periods shorter than 8 days (BL Her).
	\end{description}
\item[DCEP]		These are the classical cepheids, or Delta Cep-type
				variables. Comparatively young objects that have left the main sequence
				and evolved into the instability strip of the Hertzsprung-Russell (H-R)
				diagram, they obey the well-known Cepheid period-luminosity relation and
				belong to the young disk population. DCEP stars are present in open
				clusters. They display a certain relation between the shapes of their
				light curves and their periods.
	\begin{description}
\item[DCEPS] 	These are Delta Cep variables having light amplitudes
				\textless{}0.5 mag in V (\textless{}0.7 mag in B) and almost symmetrical
				light curves (M-m approx. 0.4 - 0.5 periods); as a rule, their periods
				do not exceed 7 days. They are probably first-overtone pulsators and/or
				are in the first transition across the instability strip after leaving
				the main sequence (SU Cas).
	
				Traditionally, both Delta Cep and W Vir stars are quite often called
				Cepheids because it is often impossible to discriminate between them on
				the basis of the light curves for periods in the range 3 - 10 days.
				However, these are distinct groups of entirely different objects in
				different evolutionary stages. One of the significant spectral
				differences between W Vir stars and Cepheids is the presence, during a
				certain phase interval, of hydrogen-line emission in the former and of
				Ca II H and K emission in the latter.
	\end{description}		
\item[DSCT]		Variables of the Delta Scuti type. These are pulsating
				variables of spectral types A0-F5 III-V displaying light amplitudes from
				0.003 to 0.9 mag in V (usually several hundredths of a magnitude) and
				periods from 0.01 to 0.2 days. The shapes of the light curves, periods,
				and amplitudes usually vary greatly. Radial as well as nonradial
				pulsations are observed. The variability of some members of this type
				appears sporadically and sometimes completely ceases, this being a
				consequence of strong amplitude modulation with the lower value of the
				amplitude not exceeding 0.001 mag in some cases. The maximum of the
				surface layer expansion does not lag behind the maximum light for more
				than 0.1 periods. DSCT stars are representatives of the galactic disk
				(flat component) and are phenomenologically close to the SX Phe
				variables.
	\begin{description}
\item[DSCTC] 	Low amplitude group of Delta Sct variables (light
				amplitude \textless{}0.1 mag in V). The majority of this type's
				representatives are stars of luminosity class V; objects of this subtype
				generally are representative of the Delta Sct variables in open
				clusters.
	\end{description}
\item[GDOR]     Gamma Doradus stars. Early type F dwarfs showing
                (multiple) periods from several tenths of a day to slightly in excess of
                one day. Amplitudes usually do not exceed 0.1 mag. Presumably low degree
                g-mode non-radial pulsators. Prototype: gamma Dor.
\item[L]		Slow irregular variables. The light variations of these
				stars show no evidence of periodicity, or any periodicity present is
				very poorly defined and appears only occasionally. Like for the type I,
				stars are often attributed to this type because of being insufficiently
				studied. Many type L variables are really semiregulars or belong to
				other types.
	\begin{description}
\item[LB]		Slow irregular variables of late spectral types (K, M, C,
				S); as a rule, they are giants (CO Cyg). This type is also ascribed, in
				the GCVS, to slow red irregular variables in the case of unknown
				spectral types and luminosities.
\item[LC]		Irregular variable supergiants of late spectral types
				having amplitudes of about 1 mag in V (TZ Cas).

\item[LPB]      The comparatively long-period pulsating B stars (periods exceeding (LBV) one day).
	\end{description}
\item[M]		Mira (Omicron) Ceti-type variables. These are long-period
				variable giants with characteristic late-type emission spectra (Me, Ce,
				Se) and light amplitudes from 2.5 to 11 mag in V. Their periodicity is
				well pronounced, and the periods lie in the range between 80 and 1000
				days. Infrared amplitudes are usually less than in the visible and may
				be \textless{}2.5 mag. For example, in the K band they usually do not
				exceed 0.9 mag. If the amplitudes exceed 1 - 1.5 mag , but it is not
				certain that the true light amplitude exceeds 2.5 mag, the symbol ``M''
				is followed by a colon, or the star is attributed to the semiregular
				class with a colon following the symbol for that type (SR).
\item[PVTEL] 	Variables of the PV Telescopii type. These are helium
				supergiant Bp stars with weak hydrogen lines and enhanced lines of He
				and C. They pulsate with periods of approximately 0.1 to 1 days, or vary
				in brightness with an amplitude of 0.1 mag in V during a time interval
				of about a year.
\item[RPHS]     Very rapidly pulsating hot (subdwarf B) stars. Typical
                periods are hundreds of seconds, amplitudes are within several
                hundredths of a magnitude. Prototype: V361 Hya = EC 14026-2647.
\item[RR]		Variables of the RR Lyrae type, which are
				radially-pulsating giant A-F stars having amplitudes from 0.2 to 2 mag
				in V. Cases of variable light-curve shapes as well as variable periods
				are known. If these changes are periodic, they are called the \indexterm{Blazhko
				effect}.
	
				Traditionally, RR Lyrae stars are sometimes called short-period Cepheids
				or cluster-type variables. The majority of these stars belong to the
				spherical component of the Galaxy; they are present, sometimes in large
				numbers, in some globular clusters, where they are known as pulsating
				horizontal-branch stars. Like Cepheids, maximum expansion velocities of
				surface layers for these stars practically coincide with maximum
				light.
	\begin{description}
\item[RR(B)] 	RR Lyrae variables showing two simultaneously operating
				pulsation modes, the fundamental tone with the period P0 and the first
				overtone, P1 (AQ Leo). The ratio P1/P0 is approximately 0.745;
\item[RRAB] 	RR Lyrae variables with asymmetric light curves (steep
				ascending branches), periods from 0.3 to 1.2 days, and amplitudes from
				0.5 to 2 mag in V;
\item[RRC]		RR Lyrae variables with nearly symmetric, sometimes
				sinusoidal, light curves, periods from 0.2 to 0.5 days, and amplitudes
				not greater than 0.8 mag in V (SX UMa).
	\end{description}
\item[RV]		Variables of the RV Tauri type. These are radially
				pulsating supergiants having spectral types F-G at maximum light and K-M
				at minimum. The light curves are characterized by the presence of double
				waves with alternating primary and secondary minima that can vary in
				depth so that primary minima may become secondary and vice versa. The
				complete light amplitude may reach 3-4 mag in V. Periods between two
				adjacent primary minima (usually called formal periods) lie in the range
				30-150 days (these are the periods appearing in the Catalogue). Two
				subtypes, RVA and RVB, are recognized:
	\begin{description}
\item[RVA]		RV Tauri variables that do not vary in mean magnitude (AC Her);
\item[RVB]		RV Tauri variables that periodically (with periods from 600
				to 1500 days and amplitudes up to 2 mag in V) vary in mean magnitude (DF
				Cyg, RV Tau).
	\end{description}
\item[SR]		Semiregular variables, which are giants or supergiants of
				intermediate and late spectral types showing noticeable periodicity in
				their light changes, accompanied or sometimes interrupted by various
				irregularities. Periods lie in the range from 20 to \textgreater{}2000
				days, while the shapes of the light curves are rather different and
				variable, and the amplitudes may be from several hundredths to several
				magnitudes (usually 1-2 mag in V).
	\begin{description}
\item[SRA]		Semiregular late-type (M, C, S or Me, Ce, Se) giants
				displaying persistent periodicity and usually small (\textless{}2.5 mag
				in V) light amplitudes (Z Aqr). Amplitudes and light-curve shapes
				generally vary and periods are in the range of 35-1200 days. Many of
				these stars differ from Miras only by showing smaller light
				amplitudes;
\item[SRB]		Semiregular late-type (M, C, S or Me, Ce, Se) giants with
				poorly defined periodicity (mean cycles in the range of 20 to 2300 days)
				or with alternating intervals of periodic and slow irregular changes,
				and even with light constancy intervals (RR CrB, AF Cyg). Every star of
				this type may usually be assigned a certain mean period (cycle), which
				is the value given in the Catalogue. In a number of cases, the
				simultaneous presence of two or more periods of light variation is
				observed;
\item[SRC]		Semiregular late-type (M, C, S or Me, Ce, Se) supergiants
				(Mu Cep) with amplitudes of about 1 mag and periods of light variation
				from 30 days to several thousand days;
\item[SRD]		Semiregular variable giants and supergiants of F, G, or K
				spectral types, sometimes with emission lines in their spectra.
				Amplitudes of light variation are in the range from 0.1 to 4 mag, and
				the range of periods is from 30 to 1100 days (SX Her, SV UMa).
\item[SRS]      Semiregular pulsating red giants with short period
                (several days to a month), probably high-overtone pulsators. Prototype: AU Ari.
	\end{description}
\item[SXPHE] 	Phenomenologically, these resemble DSCT (Delta Sct)
				variables and are pulsating subdwarfs of the spherical component, or old
				disk galactic population, with spectral types in the range A2-F5. They
				may show several simultaneous periods of oscillation, generally in the
				range 0.04-0.08 days, with variable-amplitude light changes that may
				reach 0.7 mag in V. These stars are present in globular clusters.
\item[ZZ]		ZZ Ceti variables. These are nonradially pulsating white
				dwarfs that change their brightnesses with periods from 30 s to 25 min
				and amplitudes from 0.001 to 0.2 mag in V. They usually show several
				close period values. Flares of 1 mag are sometimes observed; however,
				these may be explained by the presence of close UV Ceti companions.
	
				These variables are divided into the following subtypes:
	\begin{description}
\item[ZZA]		ZZ Cet-type variables of DA spectral type (ZZ Cet) having
				only hydrogen absorption lines in their spectra;
\item[ZZB]		ZZ Cet-type variables of DB spectral type having only
				helium absorption lines in their spectra.
\item[ZZO]		ZZ Cet-type variables of the DO spectral type showing
                HeII and and CIV absorption lines in their spectra.
	\end{description}
\end{description}

\subsubsection{Rotating Variable Stars}
\label{rotating-variable-stars}

Variable stars with nonuniform surface brightness and/or ellipsoidal
shapes, whose variability is caused by axial rotation with respect to
the observer. The nonuniformity of surface brightness distributions may
be caused by the presence of spots or by some thermal or chemical
inhomogeneity of the atmosphere caused by a magnetic field whose axis is
not coincident with the rotation axis. These stars are subdivided into
the following types:

\begin{description}
\item[ACV]		Alpha2 Canum Venaticorum variables. These are
				main-sequence stars with spectral types B8p-A7p and displaying strong
				magnetic fields. Spectra show abnormally strong lines of Si, Sr, Cr, and
				rare earths whose intensities vary with rotation. They exhibit magnetic
				field and brightness changes (periods of 0.5-160 days or more). The
				amplitudes of the brightness changes are usually within 0.01-0.1 mag in V.
	\begin{description}
\item[ACVO]		Rapidly oscillating Alpha2 CVn variables. These are
				nonradially pulsating, rotating magnetic variables of Ap spectral type
				(DO Eri). Pulsation periods are in the range of 6-12 mmag (0.004-0.01
				days), while amplitudes of light variation caused by the pulsation are
				about 0.01 mag in V. The pulsational variations are superposed on those
				caused by rotation.
	\end{description}
\item[BY]		BY Draconis-type variables, which are emission-line dwarfs
				of dKe-dMe spectral type showing quasiperiodic light changes with
				periods from a fraction of a day to 120 days and amplitudes from several
				hundredths to 0.5 mag in V. The light variability is caused by axial
				rotation of a star with a variable degree of nonuniformity of the
				surface brightness (spots) and chromospheric activity. Some of these
				stars also show flares similar to those of UV Cet stars, and in those
				cases they also belong to the latter type and are simultaneously
				considered eruptive variables.
\item[ELL]		Rotating ellipsoidal variables (b Per, Alpha Vir). These
				are close binary systems with ellipsoidal components, which change
				combined brightnesses with periods equal to those of orbital motion
				because of changes in emitting areas toward an observer, but showing no
				eclipses. Light amplitudes do not exceed 0.1 mag in V.
\item[FKCOM] 	FK Comae Berenices-type variables. These are rapidly
				rotating giants with nonuniform surface brightnesses, which have G-K
				spectral types with broad H and K Ca II emission and sometimes Halpha.
				They may also be spectroscopic binary systems. Periods of light
				variation (up to several days) are equal to rotational periods, and
				amplitudes are several tenths of a magnitude. It is not excluded that
				these objects are the product of further evolution of EW (W UMa) close
				binary systems (see below).
\item[PSR]		Optically variable pulsars (CM Tau), which are rapidly
				rotating neutron stars with strong magnetic fields, radiating in the
				radio, optical, and X-ray regions. Pulsars emit narrow beams of
				radiation, and periods of their light changes coincide with rotational
				periods (from 0.004 to 4 s), while amplitudes of the light pulses reach
				0.8 mag.
\item[R]        Close binary systems characterized by the presence of
                strong reflection (re-radiation) of the light of the hot star
                illuminating the surface of the cooler companion. Light curves are
                sinusoidal with the period equal to Porb, maximum brightness coinciding
                with the passage of the hot star in front of the companion. The eclipse
                may be absent. The range of light variation is about 0.5-1.0mag V (KV Vel).
\item[SXARI] 	SX Arietis-type variables. These are main-sequence
				B0p-B9p stars with variable-intensity He I and Si III lines and magnetic
				fields. They are sometimes called helium variables. Periods of light and
				magnetic field changes (about 1 day) coincide with rotational periods,
				while amplitudes are approximately 0.1 mag in V. These stars are
				high-temperature analogs of the ACV variables.
\end{description}

\subsubsection{Cataclysmic (Explosive and Novalike) Variables}
\label{cataclysmic-explosive-and-novalike-variables}

These are variable stars showing outbursts caused by thermonuclear burst
processes in their surface layers (novae) or deep in their interiors
(supernovae). We use the term ``novalike'' for variables that show
novalike outbursts caused by rapid energy release in the surrounding
space (UG-type stars - see below) and also for objects not displaying
outbursts but resembling explosive variables at minimum light by their
spectral (or other) characteristics. The majority of explosive and
novalike variables are close binary systems, their components having
strong mutual influence on the evolution of each star. It is often
observed that the hot dwarf component of the system is surrounded by an
accretion disk formed by matter lost by the other, cooler, and more
extended component. This category is subdivided into the following
types:

\begin{description}
\item[N]		Novae. Close binary systems with orbital periods from 0.05
				to 230 days. One of the components of these systems is a hot dwarf star
				that suddenly, during a time interval from one to several dozen or
				several hundred days, increases its brightness by 7-19 mag in V, then
				returns gradually to its former brightness over several months, years,
				or decades. Small changes at minimum light may be present. Cool
				components may be giants, subgiants, or dwarfs of K-M type. The spectra
				of novae near maximum light resemble A-F absorption spectra of luminous
				stars at first. Then broad emission lines (bands) of hydrogen, helium,
				and other elements with absorption components indicating the presence of
				a rapidly expanding envelope appear in the spectrum. As the light
				decreases, the composite spectrum begins to show forbidden lines
				characteristic of the spectra of gas nebulae excited by hot stars. At
				minimum light, the spectra of novae are generally continuous or resemble
				the spectra of Wolf-Rayet stars. Only spectra of the most massive
				systems show traces of cool components.
				
				Some novae reveal pulsations of hot components with periods of
				approximately 100 s and amplitudes of about 0.05 mag in V after an
				outburst. Some novae eventually turn out to be eclipsing systems.
				According to the features of their light variations, novae are
				subdivided into fast (NA), slow (NB), very slow (NC), and recurrent (NR)
				categories.
	\begin{description}
\item[NA]		Fast novae displaying rapid light increases and then, having
				achieved maximum light, fading by 3 mag in 100 or fewer days (GK Per);
\item[NB]		Slow novae that fade after maximum light by 3 mag in
				\textgreater{}= 150 days (RR Pic). Here the presence of the well-known
				``dip'' in the light curves of novae similar to T Aur and DQ Her is not
				taken into account: The rate of fading is estimated on the basis of a
				smooth curve, its parts before and after the ``dip'' being a direct
				continuation of one another;
				\item[NC] Novae with a very slow development and remaining at maximum
				light for more than a decade, then fading very slowly. Before an
				outburst these objects may show long-period light changes with
				amplitudes of 1-2 mag in V (RR Tel); cool components of these systems
				are probably giants or supergiants, sometimes semiregular variables, and
				even Mira variables. Outburst amplitudes may reach 10 mag. High
				excitation emission spectra resemble those of planetary nebulae,
				Wolf-Rayet stars, and symbiotic variables. The possibility that these
				objects are planetary nebulae in the process of formation is not
				excluded;
\item[NL]		Novalike variables, which are insufficiently studied objects
				resembling novae by the characteristics of their light changes or by
				spectral features. This type includes, in addition to variables showing
				novalike outbursts, objects with no bursts ever observed; the spectra of
				novalike variables resemble those of old novae, and small light changes
				resemble those typical for old novae at minimum light. However, quite
				often a detailed investigation makes it possible to reclassify some
				representatives of this highly inhomogeneous group of objects into other
				types;
\item[NR]		Recurrent novae, which differ from typical novae by the fact
				that two or more outbursts (instead of a single one) separated by 10-80
				years have been observed (T CrB).
	\end{description}
\item[SN]		Supernovae (B Cas, CM Tau). Stars that increase, as a
				result of an outburst, their brightnesses by 20 mag and more, then fade
				slowly. The spectrum during an outburst is characterized by the presence
				of very broad emission bands, their widths being several times greater
				than those of the bright bands observed in the spectra of novae. The
				expansion velocities of SN envelopes are in the thousands of km/s. The
				structure of a star after outburst alters completely. An expanding
				emission nebula results and a (not always observable) pulsar remains at
				the position of the original star. According to the light curve shape
				and the spectral features, supernovae are subdivided into types I and II.
	\begin{description}
\item[SNI]		Type I supernovae. Absorption lines of Ca II, Si, etc., but
				no hydrogen lines are present in the spectra. The expanding envelope
				almost lacks hydrogen. During 20-30 days following maximum light, the
				brightness decreases by approximately 0.1 mag per day, then the rate of
				fading slows and reaches a constant value of 0.014/day;
\item[SNII]		Type II supernovae. Lines of hydrogen and other elements
				are apparent in their spectra. The expanding envelope consists mainly of
				H and He. Light curves show greater diversity than those of type I
				supernovae. Usually after 40-100 days since maximum light, the rate of
				fading is 0.1 mag per day.
	\end{description}
\item[UG]		U Geminorum-type variables, quite often called dwarf
				novae. They are close binary systems consisting of a dwarf or subgiant
				K-M star that fills the volume of its inner Roche lobe and a white dwarf
				surrounded by an accretion disk. Orbital periods are in the range
				0.05-0.5 days. Usually only small, in some cases rapid, light
				fluctuations are observed, but from time to time the brightness of a
				system increases rapidly by several magnitudes and, after an interval of
				from several days to a month or more, returns to the original state.
				Intervals between two consecutive outbursts for a given star may vary
				greatly, but every star is characterized by a certain mean value of
				these intervals, i.e., a mean cycle that corresponds to the mean light
				amplitude. The longer the cycle, the greater the amplitude. These
				systems are frequently sources of X-ray emission. The spectrum of a
				system at minimum is continuous, with broad H and He emission lines. At
				maximum these lines almost disappear or become shallow absorption lines.
				Some of these systems are eclipsing, possibly indicating that the
				primary minimum is caused by the eclipse of a hot spot that originates
				in the accretion disk from the infall of a gaseous stream from the K-M
				star. According to the characteristics of the light changes, U Gem
				variables may be subdivided into three types: SS Cyg, SU UMa, and Z Cam.
	\begin{description}
\item[UGSS]		SS Cygni-type variables (SS Cyg, U Gem). They increase in
				brightness by 2-6 mag in V in 1-2 days and in several subsequent days
				return to their original brightnesses. The values of the cycle are in
				the range 10 days to several thousand;
\item[UGSU]		SU Ursae Majoris-type variables. These are characterized
				by the presence of two types of outbursts called ``normal'' and
				``supermaxima''. Normal, short outbursts are similar to those of UGSS
				stars, while supermaxima are brighter by 2 mag, are more than five times
				longer (wider), and occur several times less frequently. During
				supermaxima the light curves show superposed periodic oscillations
				(superhumps), their periods being close to the orbital ones and
				amplitudes being about 0.2-0.3 mag in V. Orbital periods are shorter
				than 0.1 days; companions are of dM spectral type;
\item[UGZ]		Z Camelopardalis-type stars. These also show cyclic
				outbursts, differing from UGSS variables by the fact that sometimes
				after an outburst they do not return to the original brightness, but
				during several cycles retain a magnitude between maximum and minimum.
				The values of cycles are from 10 to 40 days, while light amplitudes are
				from 2 to 5 mag in V.
	\end{description}
\item[ZAND]		Symbiotic variables of the Z Andromedae type. They are
				close binaries consisting of a hot star, a star of late type, and an
				extended envelope excited by the hot star's radiation. The combined
				brightness displays irregular variations with amplitudes up to 4 mag in
				V. A very inhomogeneous group of objects.
\end{description}

\subsubsection{Close Binary Eclipsing Systems}
\label{close-binary-eclipsing-systems}

We adopt a triple system of classifying eclipsing binary systems:
according to the shape of the combined light curve, as well as to
physical and evolutionary characteristics of their components. The
classification based on light curves is simple, traditional, and suits
the observers; the second and third classification methods take into
account positions of the binary-system components in the (MV ,B-V)
diagram and the degree of inner Roche lobe filling. Estimates are made
by applying the simple criteria proposed by Svechnikov and Istomin
(1979). The symbols for the types of eclipsing binary systems that we
use are given below.

\subsubsection{Classification based on the shape of the light curve}
\label{classification-based-on-the-shape-of-the-light-curve}

\begin{description}
\item[E]      Eclipsing binary systems. These are binary systems with
              orbital planes so close to the observer's line of sight (the inclination
              i of the orbital plane to the plane orthogonal to the line of sight is
              close to 90 deg) that the components periodically eclipse each other.
              Consequently, the observer finds changes of the apparent combined
              brightness of the system with the period coincident with that of the
              components' orbital motion.
	\begin{description}
\item[EA]     Algol (Beta Persei)-type eclipsing systems. Binaries with
              spherical or slightly ellipsoidal components. It is possible to specify,
              for their light curves, the moments of the beginning and end of the
              eclipses. Between eclipses the light remains almost constant or varies
              insignificantly because of reflection effects, slight ellipsoidality of
              components, or physical variations. Secondary minima may be absent. An
              extremely wide range of periods is observed, from 0.2 to \textgreater{}=
              10000 days. Light amplitudes are also quite different and may reach
              several magnitudes.
\item[EB]     Beta Lyrae-type eclipsing systems. These are eclipsing
              systems having ellipsoidal components and light curves for which it is
              impossible to specify the exact times of onset and end of eclipses
              because of a continuous change of a system's apparent combined
              brightness between eclipses; secondary minimum is observed in all cases,
              its depth usually being considerably smaller than that of the primary
              minimum; periods are mainly longer than 1 day. The components generally
              belong to early spectral types (B-A). Light amplitudes are usually
              \textless{}2 mag in V.
\item[EP]     Stars showing eclipses by their planets. Prototype: V0376 Peg.
\item[EW]     W Ursae Majoris-type eclipsing variables. These are
              eclipsing binaries with periods shorter than 1 days, consisting of ellipsoidal
              components almost in contact and having light curves for which it is
              impossible to specify the exact times of onset and end of eclipses. The
              depths of the primary and secondary minima are almost equal or differ
              insignificantly. Light amplitudes are usually \textless{}0.8 mag in V.
              The components generally belong to spectral types F-G and
              later.
	\end{description}
\end{description}

\subsubsection{Classification according to the components' physical characteristics}
\label{classification-according-to-the-components-physical-characteristics}

\begin{description}
\item[GS]     Systems with one or both giant and supergiant components;
              one of the components may be a main sequence star.
\item[PN]     Systems having, among their components, nuclei of planetary nebulae (UU Sge).
\item[RS]     RS Canum Venaticorum-type systems. A significant property
              of these systems is the presence in their spectra of strong Ca II H and
              K emission lines of variable intensity, indicating increased
              chromospheric activity of the solar type. These systems are also
              characterized by the presence of radio and X-ray emission. Some have
              light curves that exhibit quasi sine waves outside eclipses, with
              amplitudes and positions changing slowly with time. The presence of this
              wave (often called a distortion wave) is explained by differential
              rotation of the star, its surface being covered with groups of spots;
              the period of the rotation of a spot group is usually close to the
              period of orbital motion (period of eclipses) but still differs from it,
              which is the reason for the slow change (migration) of the phases of the
              distortion wave minimum and maximum in the mean light curve. The
              variability of the wave's amplitude (which may be up to 0.2 mag in V) is
              explained by the existence of a long-period stellar activity cycle
              similar to the 11-year solar activity cycle, during which the number and
              total area of spots on the star's surface vary.
\item[WD]     Systems with white-dwarf components.
\item[WR]     Systems having Wolf-Rayet stars among their components (V444 Cyg).
\end{description}

\subsubsection{Classification based on the degree of filling of inner Roche lobes}
\label{classification-based-on-the-degree-of-filling-of-inner-roche-lobes}

\begin{description}
\item[AR]     Detached systems of the AR Lacertae type. Both components
              are subgiants not filling their inner equipotential surfaces.
\item[D]      Detached systems, with components not filling their inner Roche lobes.
	\begin{description}
\item[DM]     Detached main-sequence systems. Both components are
              main-sequence stars and do not fill their inner Roche lobes.
\item[DS]     Detached systems with a subgiant. The subgiant also does
              not fill its inner critical surface.
\item[DW]     Systems similar to W UMa systems in physical properties (KW, see below), but not in contact.
	\end{description}
\item[K]      Contact systems, both components filling their inner critical surfaces.
	\begin{description}
\item[KE]     Contact systems of early (O-A) spectral type, both
              components being close in size to their inner critical surfaces.
\item[KW]     Contact systems of the W UMa type, with ellipsoidal
              components of F0-K spectral type. Primary components are main-sequence
              stars and secondaries lie below and to the left of the main sequence in the (MV,B-V) diagram.
	\end{description}
\item[SD]     Semidetached systems in which the surface of the less
              massive component is close to its inner Roche lobe.
\end{description}

The combination of the above three classification systems for eclipsing binaries
results in the assignment of multiple classifications for object types.
These are separated by a solidus (``/'') in the data field. Examples are:
E/DM, EA/DS/RS, EB/WR, EW/KW, etc.

\subsubsection{Optically Variable Close Binary Sources of Strong, Variable X-ray Radiation (X-ray Sources)}
\label{optically-variable-close-binary-sources-of-strong-variable-x-ray-radiation-x-ray-sources}

\begin{description}
\item[AM]    AM Her type variables; close binary systems consisting of
	         a dK-dM type dwarf and of a compact object with strong magnetic field,
	         characterized by variable linear and circular polarization of light. The
             total range of light variations may reach 4-5 mag V.
\item[X]     Close binary systems that are sources of strong, variable
             X-ray emission and which do not belong to or are not yet attributed to
             any of the above types of variable stars. One of the components of the
             system is a hot compact object (white dwarf, neutron star, or possibly a
             black hole). X-ray emission originates from the infall of matter onto
             the compact object or onto an accretion disk surrounding the compact
             object. In turn, the X-ray emission is incident upon the atmosphere of
             the cooler companion of the compact object and is reradiated in the form
             of optical high-temperature radiation (reflection effect), thus making
             that area of the cooler companion's surface an earlier spectral type.
             These effects lead to quite a peculiar complex character of optical
             variability in such systems. These objects may be subdivided into the
             following types:
	\begin{description}
\item[XB]    X-ray bursters. Close binary systems showing X-ray and
             optical bursts, their duration being from several seconds to ten
             minutes, with amplitudes of about 0.1 mag in V (V 801 Ara, V 926 Sco);
\item[XF]    Fluctuating X-ray systems showing rapid variations of X-ray
             (Cygnus X-1 = V1357 Cyg) and optical (V821 Ara) radiation on time scales of dozens of milliseconds;
\item[XI]    X-ray irregulars. Close binary systems consisting of a hot
             compact object surrounded by an accretion disk and a dA - dM-type dwarf.
             These display irregular light changes on time scales of minutes and
             hours, and amplitudes of about 1 mag in V. Superposition of a periodic
             variation because of orbital motion is possible (V818 Sco);
\item[XJ]    X-ray binaries characterized by the presence of relativistic
             jets evident at X-ray and radio wavelengths, as well as in the optical
             spectrum in the form of emission components showing periodic
             displacements with relativistic velocities (V1343 Aql);
\item[XND]   X-ray, novalike (transient) systems containing, along with
             a hot compact object, a dwarf or subgiant of G-M spectral type. These
             systems occasionally rapidly increase in brightness by 4-9 mag in V, in
             the visible simultaneously with the X-ray range, with no envelope
             ejected. The duration of the outburst may be up to several months (V616 Mon);
\item[XNG]   X-ray, novalike (transient) systems with an early-type
             supergiant or giant primary component and a hot compact object as a
             companion. Following the main component's outburst, the material ejected
             by it falls onto the compact object and causes, with a significant
             delay, the appearance of X rays. The amplitudes are about 1-2 mag in V (V725 Tau);
\item[XP]    X-ray pulsar systems. The primary component is usually an
             ellipsoidal early-type supergiant. The reflection effect is very small
             and light variability is mainly caused by the ellipsoidal primary
             component's rotation. Periods of light changes are between 1 and 10
             days; the period of the pulsar in the system is from 1 s to 100 min.
             Light amplitudes usually do not exceed several tenths of a magnitude
             (Vela X-1 = GP Vel);
\item[XPR]   X-ray pulsar systems featuring the presence of the
             reflection effect. They consist of a dB-dF-type primary and an X-ray
             pulsar, which may also be an optical pulsar. The mean light of the
             system is brightest when the primary component is irradiated by X rays;
             it is faintest during a low state of the X-ray source. The total light
             amplitude may reach 2-3 mag in V (HZ Her);
\item[XPRM,XM] X-ray systems consisting of a late-type dwarf (dK-dM)
             and a pulsar with a strong magnetic field. Matter accretion on the
             compact object's magnetic poles is accompanied by the appearance of
             variable linear and circular polarization; hence, these systems are
             sometimes known as ``polars''. The amplitudes of the light changes are
             usually about 1 mag in V but, provided that the primary component is
             irradiated by X rays, the mean brightness of a system may increase by 3
             mag in V. The total light amplitude may reach 4-5 mag in V (AM Her, AN UMa).
	\end{description}
\end{description}

If the beam of X-ray emission originating at the magnetic poles of the
rotating hot compact object does not pass through the observer's
position and the system is not observed as a pulsar, the letter ``P'' in
the above symbols for X-ray-system types is not used. If an X-ray
system is also an eclipsing or an ellipsoidal variable, the X-ray symbol
is preceded by ``E'' or ``ELL'' joined with the X-ray symbol by a ``+''
sign (e.g., E+X, ELL+X).

\subsubsection{Other Symbols}\label{other-symbols}

In addition to the variable-star types described above, certain other
symbols that need to be explained will be found in the Type data field:

\begin{description}
\item[BLLAC] Extragalactic BL Lacertae-type objects. These are
             compact quasistellar objects showing almost continuous spectra with weak
             emission and absorption lines and relatively rapid irregular light
             changes with amplitudes up to 3 mag in V or more. Sources of strong
             X-ray radiation and radio waves, their emission displays strong and
             variable linear polarization in the visible and infrared spectral
             regions. Some objects of this type, considered erroneously to be
             variable stars and designated in the GCVS system, will probably
             sometimes be included in the main table of the Catalogue in the future, too.
\item[CST]   Nonvariable stars, formerly suspected to be variable and
             hastily designated. Further observations have not confirmed their variability.
\item[GAL]   Optically variable quasistellar extragalactic objects
             (active galactic nuclei {[}AGNs{]}) considered to be variable stars by mistake.
\item[L:]    Unstudied variable stars with slow light changes.
\item[QSO]   Optically variable quasistellar extragalactic sources
             (quasars) that earlier were erroneously considered to be variable stars.
\item[S]     Unstudied variable stars with rapid light changes.
\item[*]     Unique variable stars outside the range of the
             classifications described above. These probably represent either short
             stages of transition from one variability type to another or the
             earliest and latest evolutionary stages of these types, or they are
             insufficiently studied members of future new types of variables.
\item[+]     If a variable star belongs to several types of light
             variability simultaneously, the types are joined in the Type field by a
             ``+'' sign (e.g., E+UG, UV+BY).
\item[:]     Uncertainty flag on Type of Variability
\end{description}

\section{Double Stars}
\label{sec:StarCatalogues:DoubleStars}

Since version 0.15.0\newFeature{0.15.0} Stellarium supports the subset of double stars
from WDS\footnote{\url{http://www.astro.gsu.edu/wds/} or \url{https://www.stelledoppie.it/index2.php?section=1}} which have HIP
identifiers (i.e., stars from the Hipparcos catalog). Stellarium's Catalog
of Double Stars is based on the \citetp{2001AJ....122.3466M}. %\emph{Washington Double Star Catalog} 
Cross reference data for WDS and HIP designations is based on the 
XHIP catalog \citep{2012AstL...38..331A} through the astronomical 
database SIMBAD  \citep{2000A&AS..143....9W}.

\subsection{Double Star Catalog File Format}
\label{sec:StarCatalogues:DoubleStars:format}

%\subsubsection{General Description}
%\label{sec:StarCatalogues:DoubleStars:general}

Stellarium's Double Stars catalog data is kept
in the \file{stars/hip\_gaia3} sub-directory of the Installation Directory and/or
User Directory (see section~\ref{sec:Directories}).


\subsubsection{File Format}
\label{sec:StarCatalogues:DoubleStars:file}

The \file{wds\_hip\_part.dat} file contains data about 16411
\indexterm{double stars} which have HIP identifiers and
stored in plain text format with tab delimiters.

\begin{table}[htb]
\begin{tabularx}{\textwidth}{l|c|X}\toprule
\emph{Name} & \emph{Type} & \emph{Description}\\\midrule
HIP      		& int    & HIP identifier for the star. Used for reference to the main star catalogue.\\%\midrule
WDS      		& string & WDS designation for the double star.\\%\midrule
Observation		& int	 & Date of last satisfactory observation (in years).\\%\midrule
Position Angle  & float  & Position Angle at date of last satisfactory observation (in decimal degrees).\\%\midrule
Separation 		& float  & Separation at date of last satisfactory observation (in arcsec).\\\bottomrule
\end{tabularx}
\caption{Format of \file{wds\_hip\_part.dat}}
\label{tab:StarCatalogues:DoubleStars:file}
\end{table}

\subsection{Designations}
\label{sec:StarCatalogues:DoubleStars:Designations}

In version 0.20.4\newFeature{0.20.4} we incorporated support for additional designations 
for \indexterm{double stars} which have also HIP identification. All these designations 
were added especially for double star observers and people who use Stellarium as 
reference software. The cross-identification data for double stars were obtained 
from \emph{The Cambridge Double Star Atlas}~\citep{2009cdsa.book, 2015cdsa.book.....M}.

In version 0.21.1 we updated cross-identification data according to \emph{The Washington Double Star Catalog} data~\citep{2001AJ....122.3466M}.

\begin{longtable}{l|l|c|p{59mm}|l}\toprule
\emph{Traditional ID} & \emph{Modern ID}\footnote{Latinized and capitalized version of traditional ID} & \emph{N.}\footnote{Note: * --- Stellarium uses traditional designation of double star} & \emph{Observer} & \emph{Active period} \\\midrule
--- & A   	&   & R. G. Aitken & 1891---1937 \\
--- & ABH &   & H.A. Abt & 1961---2006 \\
--- & ABT &   & G. Abetti & 1911---1925 \\
AC	& AC  &   & A. Clark & 1857---1864 \\
---	& ACA &   & A.C. Ashcraft & 2005---2007 \\
ACG	& ACG &   & \emph{Astrographic Catalog (Greenwich)} & \\
--- & ADA &   & C. Adam & 2014 \\
--- & ADM &   & E.R. Adams & 2012---2013\\
--- & AG  &   & \emph{Astronomische Gesellschaft Katalog} & 1903 \\
AGC & AGC &   & A.G. Clark & 1862 \\
--- & AHD &   & A. Ahad & 2010---2020 \\
--- & ALB &   & \emph{Albany Observatory} & 1910--1936 \\
--- & ALC &   & C. Allen & 2000---2014 \\
--- & ALD &   & H.L. Alden & 1924---1957 \\
--- & ALI &   & A. Ali & 1949---1955 \\
--- & ALL &   & R.M. Aller & 1930---1960 \\
--- & ALP &   & P.R. Allen & 2007---2012 \\
--- & AMM &   & M. Ammler-von Eiff & 2016 \\
--- & AMS &   & A. Amosse & 2012---2020 \\
--- & ANJ &   & J.A. Anderson & 1920 \\
--- & ANT &   & R. Anton & 2001--- \\
AOT & AOT &   & \emph{Astronomical Observatory Torino} & \\
--- & ARA &   & S. Aravamudan & 1961---1963 \\
Arg	& ARG &   & F.W.A. Argelander & 1857---1865 \\
--- & ARI &   & E. Aristidi & 1997---1999 \\
--- & ARN &   & D. Arnold & 2001---2013 \\
--- & ARR &   & S. Arribas & 1998 \\
--- & ARU &   & A.N. Argue & 1984---1992 \\
--- & ARY &   & R.W. Argyle & 1977---2020 \\
--- & AST &   & \emph{HST FGS Astrometry Science Team} & 1991--- \\
--- & AZC &   & A.A. Azcona & 2012---2020 \\\midrule
---	& B   &   & W.H. van den Bos & 1921---1968 \\
--- & BAG &   & Yu.Yu. Balega & 1977---2015 \\
--- & BAK &   & G.A. Bakos & 2006 \\
--- & BAL &   & R. Baillaud & 1924---1944 \\
--- & BAR &   & E.E. Barnard & 1884---1922 \\
--- & BAS &   & M.A. Barstow & 2001 \\
BAT & BAT &   & H. Batterman & \\
--- & BAU &   & C. Bauer & 1994---2007 \\
--- & BAZ &   & P. Baize & 1927---1994 \\
--- & BDK &   & \emph{Brown Dwarf Kinematic Project Catalog} & 2010 \\
--- & BDT &   & T.D. Brandt & 2014--- \\
--- & BED &   & T.R. Bedding & 1993---1994 \\
--- & BEU &   & J.-L. Beuzit & 2004 \\
--- & BEW &   & U. Bastian, J. Eisloeffel, K. Wiese & 1988 \\
--- & BGH &   & S. van den Bergh & 1958 \\
--- & BGR &   & J.-P. Berger & 2011 \\
--- & BHA &   & T.P. Bhaskaran & 1916---1938 \\
--- & BHJ &   & A. Blaauw, W.A. Hiltner, H.L. Johnson & 1959 \\
--- & BID &   & W.P. Bidelman & 1976---1980 \\
Big & BIG &   & G. Bigourdan & 1883---1903 \\
--- & BIL &   & B.A. Biller & 2006---2014 \\
--- & BKO &   & E. Berko & 2002---2012 \\
--- & BLA &   & A. Blazit & 1877---1987 \\
--- & BLK &   & D.R. Blackmore & 1990 \\
Bll & BLL &   & R.S. Ball & 1872---1884 \\
--- & BLM &   & B.L. Morgan & 1978---1982 \\
--- & BLO &   & M. Bloch & 1931 \\
--- & BNK &   & A. Brandeker & 2001---2006 \\
--- & BNS &   & D. Baines & 2006 \\
--- & BNU &   & D. Bonneau & 1979---2011 \\
--- & BNV &   & M. Bonavita & 2007---2020 \\
--- & BNW &   & W.A. Baranov & 1909 \\
--- & BOD &   & A.F. Boden & 1998---2012 \\
--- & BOT &   & G. Boettger & 1958---1967 \\
Bow & BOW &   & W.M. Bowyer & 1896---1921 \\
--- & BPM &   & \emph{Bordeaux PM2000 catalogue} & 2010 \\
--- & BRC &   & C. Briceno & 1998---2017 \\
BRD & BRD &   & F. Bird & \\
--- & BRF &   & F.L. Brown & 1916---1927 \\
--- & BRG &   & C. Bergfors & 2010---2016 \\
--- & BRH &   & W. Barth & 1994 \\
--- & BRR &   & W. Brandner & 1995---2004 \\
--- & BRT &   & S.G. Barton & 1926---1957 \\
BrsO & BSO &   & \emph{Brisbane Observatory} & 1834---1851 \\
--- & BTG &   & A. Bertoglio & 2007---2010 \\
--- & BTM &   & M. Bottom & 2015 \\
--- & BTZ &   & E. Bernewitz & 1920---1963 \\
$\beta$ & BU  & * & S. Burnham & 1873---1918 \\
$\beta$ pm & BUP & * & \emph{(Proper motion pairs)} & 1873---1918 \\
--- & BVD &   & R. Benavides & 2003---2018 \\
--- & BWL &   & B.P. Bowler & 2012---2018 \\
--- & BYV &   & V. Bailey & 2014 \\\midrule
--- & CAB &   & J.A. Caballero & 2005---2018 \\
--- & CAN &   & S.M. Caballero-Nieves & 2014---2020 \\
--- & CAR &   & M. Carbillet & 1996---1998 \\
--- & CAT &   & C. Catala & 2006 \\
--- & CBL &   & R. Caballero & 2008---2015 \\
--- & CHE &   & Stanislas Chevalier & 1908---1925 \\
--- & CHM &   & J. Chaname & 2004 \\
--- & CHN &   & M. Cohen & 1979---1983 \\
--- & CHR &   & \emph{Center for High Ang. Res. Astronomy (speckle interferometry systems)} & \\
--- & CHT &   & A.C. Cheetham & 2015---2018 \\
--- & CIA &   & \emph{Center for High Angular Resolution Astronomy (CHARA Interferometric Array)} & \\
--- & CLL &   & J.L. Comellas & 2003 \\
--- & CLO &   & L.P. Close & 1997---2014 \\
--- & CLP &   & P. Clark & 2011 \\
Cog & COG &   & W.A. Cogshall & 1896---1905 \\
Com & COM &   & G.C. Comstock & 1894---1911 \\
CorO & COO &   & \emph{Cordoba Observatory} & 1883---1920 \\
Cop & COP &   & R. Copeland & 1875---1886 \\
--- & COU &   & P. Couteau & 1952---1999 \\
CpO, CapO & CPO &   & \emph{Cape Observatory} & 1822---1938 \\
--- & CRB &   & J.-L. Carballo & 2009---2016 \\
--- & CRC &   & M. Cortes-Contreras & 2014---2017 \\
--- & CRE &   & T. Currie & 2010---2020 \\
CRI & CRI &   & F. Crifo & \\
--- & CRJ &   & J.P. Crepp & 2010---2015 \\
Cru & CRU &   & L. Cruls & 1884 \\
---	& CRZ &   & K.L. Cruz & 2007 \\
--- & CSE &   & G. Csepany & 2015---2017 \\
--- & CSI &   & Sz. Csizmadia & 2009 \\
--- & CSN &   & J.C. Carson & 2005---2013 \\
--- & CSR &   & I. Coster & 2006---2016 \\
--- & CSU &   & O. Chesneau & 2014 \\
--- & CTI &   & P.S. Conti & 1971---1980 \\
--- & CTT &   & J.-F. Courtot & 1995--- \\
--- & CTU &   & J.C. Christou & 2006---2009 \\
--- & CUD &   & K.M. Cudworth & 1973 \\
--- & CUS &   & F. Cusano & 2012 \\
--- & CUY &   & J. Cuypers & 1997---1999 \\
--- & CVE &   & Z. Cvetkovic & 2005---2020 \\
--- & CVN &   & G. Chauvin & 2002---2015 \\
--- & CVP &   & R.J. Cava and R.A. Pascal, Jr. & 2004 \\
--- & CVR &   & J. Chivers & 2011---2015 \\\midrule
D   & D   &   & H. Dembowski & 1856---1884 \\
Dawes & DA  &   & W.R. Dawes & 1831---1867 \\
--- & DAE &   & S. Daemgen & 2007---2015 \\
--- & DAL &   & J.A. Daley & 1998---2018 \\
--- & DAM &   & F. Damm & 2009--- \\
--- & DAN &   & A. Danjon & 1937---1956 \\
$\delta$ & DAW & * & B.H. Dawson &1918---1933 \\
--- & DBK &   & M. De Becker & 2004---2012 \\
--- & DCH &   & G. Duchene & 1999---2018 \\
--- & DEA &   & N.R. Deacon & 2005---2020 \\
--- & DEL &   & X. Delfosse & 1999---2000 \\
--- & DIC &   & J. Dick & 1921---1963 \\
--- & DJU &   & P. Djurkovic & 1953---1973 \\
Dob & DOB &   & W.A. Doberck & 1875---1934 \\
--- & DOM &   & J. Dommanget & 1949---1979 \\
--- & DON &   & H.F. Donner & 1937---1964 \\
--- & DOO &   & E. Doolittle & 1901---1923 \\
--- & DOR &   & \emph{Dorpat Observatory} & \\
--- & DRD &   & P. Durand & 1983---2011 \\
--- & DRS &   & R.J. De Rosa & 2011---2019 \\
--- & DSG &   & \emph{DSSI-Gemini} & 2012 \\
$\Delta$ & DUN & * & J. Dunlop & 1825---1846 \\
--- & DUQ &   & A. Duquennoy & 1987---1996 \\
--- & DVG &   & M.J. del Valle Gonzalez & 2015--- \\
DYR & DYR &   & E.R. Dyer, Jr. & \\\midrule
--- & EBE &   & J. Ebersberger & 1979---1986 \\
EDG & EDG &   & D.W. Edgecomb & \\
Egb & EGB &   & H.V. Egbert & 1879---1906 \\
--- & EGG &   & O.J. Eggen & 1945---1982 \\
--- & EGN &   & P. Eggenberger & 2007---2008 \\
--- & EHR &   & D. Ehrenreich & 2010 \\
--- & EIS &   & T. Eisenbeiss & 2007 \\
--- & ELP &   & P. Elliott & 2015---2016 \\
Eng & ENG &   & R. Engelmann & 1865---1887 \\
Enh & ENH &   & B. von Engelhardt & 1887---1895 \\
--- & ENO &   & \emph{European Network Observations} & \\
--- & ES  &   & T.E. Espin & 1901---1933 \\
--- & ESG &   & S. Els & 2001 \\
--- & EVS &   & N.R. Evans & 1993---2020 \\
--- & EVT &   & T.M. Evans & 2012 \\\midrule
--- & FAB &   & C. Fabricius & 2000 \\
--- & FAL &   & J.L. Falin and F. Mignard & 1999 \\
--- & FAR &   & J. Farihi & 2004---2013 \\
--- & FGL &   & E.D. Feigelson & 1996---2006 \\
--- & FGS &   & \emph{HST Fine Guide Sensor} & 2015 \\
$\phi$ & FIN &   & W.S. Finsen & 1926---1978 \\
--- & FLE &   & J. Fleckenstein & 1956---1958 \\
--- & FMR &   & F.M. Rica Romero & 1996--- \\
--- & FOR &   & L. Forgeron & 1911 \\
--- & FOX &   & P. Fox & 1909---1946 \\
--- & FRH &   & R. Furuhjelm & 1932 \\
--- & FRK &   & W.S. Franks & 1914---1919 \\
--- & FRT &   & W.J. Forrest & 1988---1990 \\
--- & FYM &   & M. Fay & 2013---2017 \\\midrule
--- & GAA &   & A. Gallenne & 2014---2019 \\
--- & GAL &   & J. Gallo & 1912---1914 \\
GAn & GAN &   & G. Anderson & 1888---1897 \\
--- & GAT &   & G.D. Gatewood & 1973---2019 \\
--- & GAU &   & P.L. Gauchet & 1923---1928 \\
--- & GDM &   & B. Goldman & 2010 \\
--- & GHE &   & A.M. Ghez & 1991---2016 \\
--- & GIC &   & H. Giclas & 1961---1978 \\
--- & GII &   & R. Gili & 1991---2020 \\
--- & GIR &   & P.M. Girard & 1992---1994 \\
--- & GIZ &   & J.E. Gizis & 2000---2006 \\
--- & GJ  &   & W. Gliese and H. Jahreiss & 1988 \\
--- & GKI &   & D.A. Golimowski & 1993---2004 \\
Gale & GLE &   & W.F. Gale & 1897 \\
Gli & GLI &   & J.M. Gilliss & 1850---1868 \\
Glp & GLP &   & S. Glasenapp & 1888---1899 \\
--- & GMC &   & G. Micello & 2011---2019 \\
--- & GNS &   & C. Ginski & 2012---2016 \\
GNT & GNT &   & Grant & \\
--- & GOE &   & K.-D. Goecking & 1994 \\
GOL & GOL &   & H. Goldschmidt & \\
--- & GRB &   & S. Groombridge & 1913 \\
--- & GRE &   & J. Gregorio-Hetem & 1992 \\
--- & GRF &   & R.F. Griffin & 1975---2019 \\
--- & GRV &   & J. Greaves & 2004--- \\
--- & GRY &   & C.A. Grady & 2004 \\
--- & GSH &   & J. Glaisher & 1906---1908 \\
--- & GTB &   & K. Gottlieb & 1948 \\
--- & GUE &   & E.W. Guenther & 2001---2005 \\
--- & GUI &   & J. Guillaume & 1912---1931 \\
--- & GUL &   & K. Gullikson & 2016 \\
--- & GVT &   & \emph{GRAVITY collaboration} & 2018--- \\
--- & GWP &   & \emph{Garraf Wide Pairs} & 2012---2016 \\
--- & GYL &   & A.N. Goyal & 1958---1970 \\
--- & GZA &   & B. Gauza & 2012---2015 \\\midrule
H I to VI, H N & H 1 to 6, H N &   & W. Herschel & 1777---1821 \\
--- & HAT &   & P. Hartigan & 1994---2003 \\
--- & HAU &   & H. Haupt & 1964---1966 \\
Hd, HdO & HDO &   & \emph{Harvard Observatory} & 1866---1911 \\
--- & HDS &   & \emph{Hipparcos Double Star Catalog} & c. 1991 \\
--- & HEG &   & E.K. Hege & 1983---1991 \\
--- & HEI &   & W.D. Heintz & 1954---2001 \\
--- & HEL &   & K. Helminiak & 2009---2016 \\
--- & HER &   & G.H. Herbig & 1950---1998 \\
h    & HJ  &   & J.F.W. Herschel & 1826---1870 \\
--- & HJL &   & J.L. Halbwachs & 1981---2020 \\
Hl  & HL  &   & A. Hall & 1866---1894 \\
--- & HLA &   & L.A. Hirsch & 2019 \\
Hld & HLD &   & E.S. Holden & 1874---1884 \\
--- & HLG &   & G. Hill & 1970---1988 \\
--- & HLM &   & E. Holmes & 1906 \\
--- & HLN &   & F. Holden & 1963---1979 \\
--- & HMM &   & C.A. Hummel & 2001---2017 \\
--- & HNK &   & S. Hinkley & 2010---2015 \\
Ho  & HO  &   & G.W. Hough & 1883---1907 \\
--- & HOM &   & F. Hormuth & 2007 \\
Hrg & HRG &   & L. Hargrave & 1871---1892 \\
--- & HRS &   & D.L. Harris & 1947 \\
--- & HSL &   & J.N. Heasley & 2002 \\
--- & HSW &   & R. Harshaw & 1990---1993 \\
HTG & HTG &   & C.S. Hastings & \\
Hu  & HU  &   & W.J. Hussey & 1896---1914 \\
--- & HUB &   & S. Hubrig & 2001 \\
--- & HUE &   & N. Huelamo & 2009---2015 \\
--- & HWD &   & A.W. Howard & 2010 \\
Howe & HWE &   & H.A. Howe & 1876---1882 \\
--- & HZE &   & A. Heinze & 2009---2010 \\
--- & HZG &   & E. Hertzsprung & 1907---1958 \\\midrule
--- & I   &   & R.T.A. Innes & 1895---1948 \\
--- & INN &   & P.A. Ianna & 1974---1988 \\
--- & IOT &   & \emph{IOTA Array} & 1995---2011 \\
--- & ISO &   & S. Isobe & 1990---1992 \\
--- & ITF &   & I.T. Falcon & 2011---2020 \\
--- & ITO &   & Y. Itoh & 1999---2011 \\
--- & ISM &   & R.M. Ismailov & 1998---2020 \\\midrule
--- & J   &   & R. Jonckheere & 1908---1958 \\
--- & JAY &   & R. Jayawardhana & 2001---2006 \\
Jc  & JC  &   & W.S. Jacob & 1847---1860 \\
JCT & JCT &   & J.-C. Thorel & \\
--- & JEF &   & H.M. Jeffers & 1945---1978 \\
--- & JKA &   & K.A. Janes & 2017 \\
JKS & JKS &   & M.P. Jenniskens & \\
--- & JNN &   & M. Janson & 2006---2020 \\
--- & JOD &   & E. Jodar & 2013 \\
--- & JOH &   & E.L. Johnson & 1916---1948 \\
--- & JOY &   & A.H. Joy & 1941---1949 \\
--- & JRN &   & A. Jorissen & 1988 \\
--- & JSP &   & M.K. Jessup & 1937---1964 \\\midrule
--- & KAL &   & P. Kalas & 2008 \\
--- & KAM &   & P. van de Kamp & 1929---1969 \\
--- & KAR &   & M. Karovska & 1986---1997 \\
--- & KEL &   & A. Kellerer & 2007 \\
--- & KIR &   & J.D. Kirkpatrick & 1995---2014 \\
--- & KNG &   & B. Koenig & 2002---2005 \\
--- & KNP &   & G.F.G. Knipe & 1959---1969 \\
Knott & KNT &   & G. Knott & 1865---1881 \\
Ko  & KO  &   & \emph{Koenigstuhl (Max Planck Institute)} & \\
--- & KOE &   & L. Koechlin & 1979---1983 \\
--- & KOH &   & R. Koehler & 1998---2016 \\
--- & KOI &   & \emph{Kepler Object of Interest} & 2011---2018 \\
--- & KON &   & Q.M. Konopacky & 2007---2010 \\
--- & KOU &   & M.B.N. Kouwenhoven & 2005 \\
--- & KPP &   & W. Knapp & 2015--- \\
--- & KPR &   & K.W. Kamper & 1976---1992 \\
--- & KR  &   & A. Kruger & \\
KRO & KRO &   & G. Kron & \\
--- & KRS &   & C.D. Koresko & 1991---2002 \\
--- & KRU &   & E.C. Kruger & 1936---1937 \\
--- & KRV &   & P. Kervella & 2009---2019 \\
--- & KRW &   & Carlos A. Krawczenko & 2009---2020 \\
--- & KSS &   & S. Kraus & 2007---2012 \\
--- & KST &   & J.H. Kastner & 2012 \\
Ku  & KU  &   & F. Kustner & 1884---1902 \\
--- & KUI &   & G.P. Kuiper & 1929---1961 \\
Kza & KZA &   & S.M. Kazeza & 1984 \\\midrule
L   & L   &   & T. Lewis & 1891---1921 \\
--- & LAB &   & A. Labeyrie & 1974 \\
--- & LAD &   & T. Ladanyi & 2003---2007 \\
--- & LAF &   & D. Lafreniere & 2007---2014 \\
Lal & LAL &   & J.-J. de Lalande & 1825---1877 \\
Lam & LAM &   & J. Lamont & 1849 \\
--- & LAU &   & H.E. Lau & 1905---1917 \\
--- & LAW &   & N.M. Law & 2006---2010 \\
--- & LBZ &   & P. Labitzke & 1925---1929 \\
Lcl & LCL &   & N.L. Lacaille & 1750 \\
--- & LCS &   & P.W. Lucas & 2005 \\
--- & LDS, LPM &   & \emph{Luyten p.m. catalog} & \\
--- & LEC &   & J. Lecacheux & 2000---2007 \\
LEE & LEE &   & O.J. Lee & \\
--- & LEI &   & Ch. Leinert & 1986---2001 \\
--- & LEO &   & F.C. Leonard & 1918---1943 \\
--- & LEP &   & S. Lepine & 2001---2011 \\
--- & LFR &   & L. Ferrero & 2009 \\
--- & LGR &   & A.-M. Lagrange & 2006---2009 \\
--- & LI  &   & W. Li & 1994 \\
--- & LIC &   & G. Li Causi & 2013 \\
--- & LIM &   & M.C. Liu & 1996---2012 \\
--- & LIN &   & J.F. Ling & 1983---2020 \\
--- & LLO &   & J.P. Lloyd & 2002---2007 \\
--- & LMP &   & P. Lampens & 2001---2016 \\
--- & LNA &   & A. Luna & 2020 \\
--- & LOC &   & G. Locatelli & 2013---2019 \\
--- & LOS &   & F. Losse & 2005---2010 \\
--- & LOU &   & N.P. Loutrel & 2011 \\
--- & LPL &   & N. La Palombara & 2012 \\
LpO & LPO &   & \emph{La Plata Observatory} & \\
--- & LSC &   & \emph{Lowell-Southern Connecticut Collaboration} & \\
--- & LUH &   & K.L. Luhman & 2002---2013 \\
--- & LUY &   & W.J. Luyten & 1933---1969 \\
Lv  & LV  &   & F.P. Leavenworth & 1887---1930 \\
--- & LWR &   & P.J. Lowrance & 1999---2005 \\
--- & LYS &   & L. Louys & 1987 \\\midrule
Mad & MAD &   & J.H. Maedler & 1835---1859 \\
--- & MAJ &   & J. Manfroid & 1987 \\
--- & MAM &   & E.E. Mamajek & 2010---2017 \\
--- & MAR &   & C. Marois & 2008---2010 \\
--- & MBA &   & B.A. Macintosh & 2001---2003 \\
--- & MCA &   & H.A. McAlister & 1976---1996 \\
--- & MCD &   & M. Michaud & 2012 \\
--- & MCN &   & L. Maccarini & 2015---2019 \\
--- & MCT &   & C. McCarthy & 2001 \\
--- & MCY &   & D.W. McCarthy & 1982---1993 \\
--- & MEL &   & E.L. Martin & 1998---2006 \\
--- & MET &   & S.A. Metchev & 2003---2009 \\
--- & MFL &   & F.L. Martinez & 2018---2019 \\
--- & MGN &   & G.N. Mace & 2013 \\
--- & MIL &   & J.A. Miller & 1903---1907 \\
--- & MIU &   & N. Miura & 1992---1995 \\
--- & MKR &   & V.V. Makarov & 2005---2010 \\
--- & MKT &   & \emph{Mark III Interferometer} & 1988---2006 \\
--- & MLB &   & W. Milburn & 1918---1939 \\
--- & MLD &   & A. Meilland & 2008---2011 \\
MlF & MLF &   & F. Muller & 1889 \\
MlbO & MLO &   & \emph{Melbourne Observatory} & 1873---1898 \\
--- & MLR &   & P. Muller & 1938---1997 \\
--- & MOC &   & C. Moutou & 2017 \\
--- & MRI &   & E.R. Masa Martin & 2005---2015 \\
--- & MRN &   & L. Marion & 2014 \\
--- & MRZ &   & W. Markowitz & 1956---1969 \\
--- & MSH &   & T. Meshkat & 2015 \\
--- & MUA &   & A. Muller & 2011 \\
--- & MUG &   & M. Mugrauer & 2004---2020 \\
--- & MWT &   & D. Mawet & 2011---2015 \\
--- & MZA &   & J. Maiz-Apellaniz & 2007---2020 \\\midrule
--- & NAJ &   & T. Nakajima & 1989---1995 \\
--- & NHR &   & R. Neuhauser & 2000---2011 \\
--- & NLS &   & E.L. Nielsen & 2010---2016 \\
--- & NOI &   & \emph{Navy Prototype/Precision Optical Interferometer} & 1997---2019 \\
--- & NOR &   & J.R. North & 2007 \\
--- & NOV &   & B. Novakovic & 2005---2008 \\
--- & NSE &   & P. Nisenson & 1985 \\
--- & NSN &   & J. Nanson & 2013---2018 \\
--- & NVL &   & I. Novalbos & 2009---2016 \\
--- & NYS &   & O. Nys & 1963---1983 \\
NZO & NZO &   & \emph{New Zealand Observatory} & 1911 \\\midrule
--- & OCC &   & \emph{Occult. Doubles Cat.} & 1971--- \\
--- & ODE &   & M. Odenkirchen & 1995---1997 \\
--- & OL  &   & C.P. Olivier & 1907---1957 \\
--- & OLE &   & D. Olevic & 1970---2006 \\
--- & ONL &   & \emph{Occult. News Letters} & \\
--- & OPI &   & E. Opik & 1927---1932 \\
--- & ORL &   & V.G. Orlov & 2007--- \\
--- & OSO &   & M.R. Zapatero Osorio & 2004---2010 \\
--- & OSV &   & V. Osvalds & 1957---1964 \\
--- & OSW &   & T.D. Oswalt & 1988---2012 \\
--- & OUD &   & R.D. Oudmaijer & 2010 \\\midrule
--- & PAR &   & J.A. Parkhurst & 1932 \\
--- & PAT &   & J. Patience & 1998---2008 \\
--- & PBL &   & T. Pribulla & 2006---2018 \\
--- & PBS &   & T. Preibisch & 1999---2014 \\
Pck & PCK &   & E. Pickering & 1879---1912 \\
--- & PEA &   & F.G. Pease & 1925---1927 \\
Per & PER &   & J. Perrotin & 1883---1887 \\
--- & PES &   & S. Peretti & 2019 \\
--- & PET &   & D. Peter & 2012 \\
--- & PIC &   & W.H. Pickering & 1900---1937 \\
--- & PIN &   & D.J. Pinfield & 2006---2012 \\
--- & PKO &   & \emph{Pulkova Observatory} & 1965--- \\
--- & PLQ &   & E. Paloque & 1923---1939 \\
--- & PLT &   & L. Plaut & 1934---1940 \\
--- & PNT &   & L.R. de Pont & 2012--- \\
Pol & POL &   & J.A. Pollock & 1887 \\
--- & POP &   & G.M. Popovic & 1967---2007 \\
--- & POT &   & D. Potter & 2002 \\
--- & POU &   & M.A. Pourteau & 1933 \\
--- & PRB &   & R.G. Probst & 2006 \\
Prn & PRN &   & C.D. Perrine & 1899---1900 \\
PrO & PRO &   & \emph{Perth Observatory} & 1913 \\
--- & PRU &   & J.-L. Prieur & 2001---2017 \\
Pry & PRY &   & J.J.M. Perry & 1880---1886 \\
--- & PRZ &   & E. Przbyllok & 1908---1926 \\
--- & PSF &   & \emph{Project Sixteen Forty} & \\
--- & PTI &   & \emph{Palomar Testbed Interferometer} & 2001---2004 \\
--- & PTR &   & M.G. Petr & 1998 \\
Pwl & PWL &   & E.B. Powell & 1854---1892 \\
--- & PWS &   & T. Pauwels and P. Lampens & 1993 \\
Pz  & PZ  &   & G. Piazzi & 1814 \\\midrule
R   & R   &   & H.C. Russell & 1871---1892 \\
--- & RAG &   & D. Raghavan & 2004---2010 \\
--- & RAO &   & \emph{Robo AO project} & 2014---2020 \\
--- & RAS &   & D.A. Rastegaev & 2007---2014 \\
--- & RBR &   & L.C. Roberts, Jr. & 2002---2018 \\
--- & RBT &   & J.G. Robertson & 1991---1999 \\
--- & RCH &   & A. Richichi & 1994---2020 \\
--- & RDR &   & D.R. Rodriguez & 2011---2015 \\
--- & RED &   & I.N. Reid & 1982---2008 \\
REI & REI &   & K. Reinmuth & \\
--- & REP &   & B. Reipurth & 1986---2015 \\
--- & REU &   & D. Reuyl & 1931---1945 \\
--- & RGS &   & T.J. Rodigas & 2014---2016 \\
Rhd & RHD &   & Father Richaud & 1729 \\
--- & RIZ &   & A.C. Rizzuto & 2013---2020 \\
--- & RJJ &   & J. Rajagopal & 2010 \\
Rmk & RMK &   & C. Rumker & 1825---1837 \\
--- & RMU &   & J. Rameau & 2013 \\
--- & ROE &   & E.D. Roe & 1908---1919 \\
--- & ROL &   & T. Roell & 2012 \\
--- & RSS &   & J.M. Rousseau & 1996 \\
--- & RST &   & R.A. Rossiter & 1933---1964 \\
--- & RUC &   & S.M. Rucinski & 2007---2010 \\\midrule
S   & S   &   & J. South & 1826---1825 \\
--- & SAN &   & R.F. Sanford & 1921---1940 \\
--- & SAO &   & \emph{Special Astrophysical Observatory} & 2017 \\
--- & SBK &   & G.M. Seabroke & 1881---1910 \\
--- & SBN &   & E. Serabyn & 2006---2010 \\
--- & SCA &   & M. Scardia & 1979---2020 \\
--- & SCE &   & A. Schneider & 2011 \\
--- & SCJ &   & H. Schjellerup & 1906 \\
--- & SCO &   & M. Scholler & 1996---2020 \\
--- & SCP &   & C. Schnupp & 2010 \\
--- & SCZ &   & A.B. Schultz & 1998 \\
--- & SDK &   & N. Sanduleak & 1990 \\
--- & SDR &   & W.L. Sanders & 1966 \\
--- & SDT &   & T.O.B. Schmidt & 2008---2013 \\
Se  & SE  &   & A. Secchi & 1855---1865 \\
$\lambda$ & SEE & *  & T.J.J. See & 1892---1900 \\
--- & SEI &   & J. Scheiner & 1908 \\
Shb & SHB &   & J.M. Schaeberle & 1893---1897 \\
Sh  & SHJ &   & J. South and J. Herschel & 1821---1840 \\
--- & SHN &   & G. Schneider & 1993---2004 \\
--- & SHT &   & N.I. Shatsky & 1998---2014 \\
--- & SHY &   & E.J. Shaya & 2011 \\
--- & SIN &   & D. Sinachopoulos & 1988---2007 \\
--- & SKF &   & B.A. Skiff & 2000--- \\
Ski & SKI &   & A.N. Skinner & 1872---1911 \\
--- & SLE &   & G. Soulie & 1966---2007 \\
Slr & SLR &   & R.P. Sellors & 1892---1910 \\
--- & SMA &   & W.M. Smart & 1931---1932 \\
--- & SMK &   & P.C. Schmidtke & 1984---2011 \\
--- & SML &   & L.C. Smith & 2015 \\
--- & SMR &   & J. Schlimmer & 2007--- \\
--- & SMW &   & G.V. Simonov & 1937---1951 \\
Smyth & SMY &   & W.H. Smyth & 1833---1860 \\
--- & SNA &   & H. Sana & 2006---2014 \\
--- & SOZ &   & R.-D. Scholz & 2003---2017 \\
Sp  & SP  &   & J.V. Schiaparelli & 1876---1909 \\
--- & SPM &   & \emph{Springbook Middle School} & 2021 \\
--- & SRT &   & R. Shuart & 2005 \\
--- & SRW &   & R.W. Sinnott & 1999 \\
--- & SS  &   & S. Sharpless & 1954 \\
--- & ST  &   & C.L. Stearns & 1936---1939 \\
--- & STA &   & G.A. Starikova & 1962---1985 \\
--- & STE &   & C.B. Stephenson & 1965---1988 \\
$\Sigma$ & STF & *  & F. Wilhelm von Struve & 1814---1843 \\
$\Sigma$ I, $\Sigma$ II & STFA, STFB & *  & \emph{(Supplemental catalogs)} & 1814---1843 \\
--- & STG &   & G. Struve & 1932---1963 \\
--- & STH &   & H. Struve & 1901---1913 \\
--- & STI &   & J. Stein & 1926 \\
--- & STJ &   & J. Serot & 2009---2019 \\
Stone & STN &   & O. Stone & 1867---1882 \\
O$\Sigma$ & STT & *  & O. von Struve & 1832---1878 \\
O$\Sigma\Sigma$ & STTA & *  & \emph{(Supplemental catalogs)} & 1832---1878 \\
--- & STU &   & K. Sturdy & 1984---1994 \\
--- & STZ &   & M.F. Sterzik & 1997---2005 \\
--- & SUB &   & J.P. Subasavage & 2005---2017 \\
--- & SUN &   & G. Sun & 2020 \\
SWI & SWI &   & L. Swift & \\
--- & SWR &   & J. Stock & 1972---1997 \\
SyO & SYO &   & \emph{Sydney Observatory} & 1927---1986 \\
--- & SZA &   & T.B. Souza & 2020 \\\midrule
--- & TAM &   & V.S. Tamazian & 1999---2008 \\
Tar & TAR &   & K.J. Tarrant & 1886---1893 \\
--- & TDS, TDT &   & \emph{Tycho Double Star} & \\
--- & THA &   & C. Thalmann & 2009 \\
--- & THB &   & F. Thiebaut & 1995 \\
--- & THP &   & B. Thompson & 2006---2014 \\
--- & TKA &   & V.V. Tel'nyuk-Adamchuk & 1966 \\
--- & TNG &   & W.J. Tango & 1979---2009 \\
--- & TNN &   & A.M. Tanner & 2010 \\
--- & TOB &   & T. Tobal  & 2000---2016 \\
--- & TOI &   & \emph{TOI Numbers} from Stassun et al. & 2010 \\
--- & TOK &   & A.A. Tokovinin & 1979--- \\
--- & TOR &   & \emph{Torino Observatory} & 1978---1992 \\
--- & TRA &   & V. Trajkovska & 1999 \\
--- & TRN &   & N.H. Turner & 1999---2010 \\
--- & TRR &   & G. Torres & 1985---2020 \\
Tuc & TUC &   & R.H. Tucker & 1893 \\
--- & TVB &   & T.V. Bryant & 2014---2019 \\\midrule
--- & UC  &   & \emph{USNO CCD Astrographic Catalog (UCAC1, UCAC2, UCAC3, UCAC4) } & 2013---2015 \\
Upt & UPT &   & W. Upton & 1877---1878 \\
--- & UR  &   & \emph{USNO Robotic Astrometric Telescope (URAT) } & 2015---2016 \\
--- & USD &   & \emph{US Numbers} from Esplin et al. & 2005---2018 \\\midrule
--- & VAS &   & G. Vaskuti & 2011---2012 \\
--- & VBS &   & G. Van Biesbroeck & 1904---1965 \\
--- & VDK &   & J. Vanderdonck & 1909---1944 \\
--- & VGT &   & N. Vogt & 2012---2015 \\
--- & VIG &   & A. Vigan & 2012---2016 \\
--- & VKI &   & E. Vilkki & 1984---1989 \\
--- & VLM &   & W. Vollmann & 2006---2016 \\
--- & VOU &   & J.G. Voute & 1908---1956 \\
--- & VSP &   & \emph{VLT + SPHERE} & 2019--- \\
VVO & VVO &   & \emph{Van Vleck Observatory} & 1962---1993 \\
--- & VVV &   & \emph{Vista Variables in Via Lactea} & \\
--- & VYS &   & A.N. Vyssotsky & 1927---1956 \\\midrule
--- & WAK &   & R.L. Walker & 1966---1985 \\
--- & WAL &   & A. Wallenquist & 1934---1958 \\
--- & WCK &   & W.C. Wickes & 1973---1975 \\
--- & WDK &   & K. Patience Ward-Duong & 2015 \\
Webb & WEB &   & T.W. Webb & \\
--- & WEI &   & M. Weisse & 1906 \\
--- & WFC &   & \emph{Washington Fundamental Cat. - astrographs} & 1939---1999 \\
--- & WG  &   & R.W. Wrigley & 1911 \\
--- & WGA &   & G.A.H. Walker & 1994 \\
--- & WGK &   & K. Wagner & 2016---2018 \\
--- & WGT &   & G.P. Weigelt & 1978---2002 \\
WHC & WHC &   & H.C. Wilson & 1890 \\
--- & WHI &   & G.L. White & 1991---2019 \\
--- & WIL &   & J.C. Wilson & 2001 \\
--- & WIR &   & C.A. Wirtanen & 1941---1981 \\
--- & WIS &   & \emph{Wide-field Infrared Survey Explorer} & 2010---2016 \\
--- & WLF &   & Wolf & 1913 \\
Wnc & WNC &   & F.A. Winnecke & 1855---1909 \\
WNO & WNO &   & \emph{Washington Naval Observatory} & 1866---1896 \\
--- & WOR &   & C.E. Worley & 1956---2001 \\
Wrd & WRD &   & I.W. Ward & 1876---1878 \\
--- & WRH &   & R.H. Wilson & 1933---1982 \\
--- & WRS &   & K. Wierzchos & 2017---2019 \\
WS  & WS  &   & J.M. Wilson & 1875---1877 \\
--- & WSI &   & \emph{Washington Speckle Interferometer} & 1997--- \\
--- & WTT &   & J.M. Wittrock & 2016---2017 \\
--- & WZ  &   & C. Wirtz & 1912---1923 \\\midrule
--- & XMI &   & X. Miret & 2007---2016 \\\midrule
Yng & YNG &   & C.A. Young & 1881---1887 \\
--- & YR  &   & \emph{Yale-Rochester Collaboration} & \\
--- & YSC &   & \emph{Yale-Southern Connecticut  Collaboration} & \\
--- & YSJ  &   & \emph{Yale-San Juan Collaboration} & \\\midrule
--- & ZHN &   & Z.H. Zhang & 2010---2013 \\
---  & ZUC  &   & B. Zuckerman & 1992---2001 \\
---  & ZUL  &   & D.J. Zulevic & 1967---2000 \\\bottomrule
\caption{Additional designations for double stars}
\label{tab:StarCatalogues:DoubleStars:Designations}
\end{longtable}

\section{Cross-Identification Data}
\label{sec:StarCatalogues:CrossIdentificationData}

Stellarium supports\newFeature{0.14.0} cross-identification between stars from 
Hipparcos (HIP)~\citep{2012AstL...38..331A}, 
Smithsonian Astrophysical Observatory (SAO) Star Catalog J2000.0~\citep{1995yCat.1131....0S}, Henry Draper Catalogue and Extension (HD)~\citep{1993yCat.3135....0C} and 
\citetp{1964cbs..book.....H}  (BSC5 or HR) % $5^{th}$ Revised Ed. of Bright Star Catalog (Hoffleit+, 1991; HR)
catalogues.

\subsection{Cross-Identification Catalog File Format}
\label{sec:StarCatalogues:CrossIdentificationData:format}

%\subsubsection{General Description}
%\label{sec:StarCatalogues:CrossIdentificationData:general}

Stellarium's Cross-Identification catalog data is kept
in the \file{stars/hip\_gaia3} sub-directory of the Installation Directory and/or User Directory (see section~\ref{sec:Directories}).

\subsubsection{File Format}
\label{sec:StarCatalogues:CrossIdentificationData:file}

The file \file{cross-id.cat} contains cross reference data for catalogues HIP, SAO, HD and HR, which are stored in a binary format.

\begin{table}[htb]
\begin{tabularx}{\textwidth}{l|c|X}\toprule
\emph{Name} & \emph{Type}   & \emph{Description}\\\midrule
HIP         & uint64\_t     & HIP or Gaia identifier for the star. Used for reference to the main star catalogue.\\
Component   & unsigned char & Component of the star in the multiple star system.\\
SAO    	    & int           & SAO identifier for the star.\\
HD          & int           & HD identifier for the star.\\
HR          & int           & HR identifier for the star.\\\bottomrule
\end{tabularx}
\caption{Format of \file{cross-id.cat}}
\label{tab:StarCatalogues:CrossIdentificationData:file}
\end{table}


\section{Binary Star System Orbital Parameters}
\label{sec:StarCatalogues:BinaryOrbitalParameters}

Since version 25.1\newFeature{25.1} Stellarium supports modelling orbital motion of a handful of binary stars. These include

\begin{itemize}
	\item $\alpha$ Centauri A \& B
	\item Sirius A \& B
	\item 61 Cygni A \& B
	\item $\gamma$ Virginis A \& B
	\item Alula Australis A \& B
	\item Achird A \& B
	\item $\varepsilon$1 Lyr A \& B
	\item $\varepsilon$2 Lyr A \& B
	\item $\beta$ Mus A \& B
	\item Dalim A \& B
	\item $\psi$ Vel A \& B
	\item $\rho$ Oph A \& B
	\item $\zeta$ Her A \& B
	\item $\eta$ Oph A \& B
\end{itemize}

\subsection{Binary Star System Orbital Parameters Catalog File Format}
\label{sec:StarCatalogues:BinaryOrbitalParameters:format}

Stellarium's Binary Star System Orbital Parameters catalog data is kept
in the \file{stars/hip\_gaia3} sub-directory of the Installation Directory and/or User Directory (see section~\ref{sec:Directories}).

\subsubsection{File Format}
\label{sec:StarCatalogues:BinaryOrbitalParameters:file}

The file \file{binary\_orbitparam.dat} contains binary system orbital parameter data, which are stored in 
plain text format (\ref{tab:StarCatalogues:BinaryOrbitalParameters:file}).

\begin{table}[htb]
	\begin{tabularx}{\textwidth}{l|c|X}\toprule
	  \emph{Name}                    & \emph{Type} & \emph{Description}                     \\\midrule
	  binary period  $p$             & double      & Orbital period [days]                  \\%\midrule
	  eccentricity   $e$             & float       & Eccentricity                           \\%\midrule
	  inclination    $i$             & float       & Orbit inclination [rad]                \\%\midrule
	  big omega      $\Omega$        & float       & Position angle of ascending node [rad] \\%\midrule
	  small omega    $\omega$        & float       & Argument of periastron [rad]           \\%\midrule
	  periastron epoch $JD$          & double      & Julian Date at periastron passage      \\%\midrule
	  semi major     $a$             & double      & Angular semi-major axis [arcsec]       \\%\midrule
	  bary distance  $\Delta$        & double      & distance [parsec]                      \\%\midrule
	  data epoch     $E$             & double      & Julian Date of the data                \\%\midrule
	  bary ra        $\alpha$        & double      & Right Ascension of the barycenter [rad]\\%\midrule
	  bary dec       $\delta$        & double      & Declination of the barycenter  [rad]   \\%\midrule
	  bary rv        $RV$            & double      & Barycenter radial velocity [km/s]      \\%\midrule
	  primary mass   $m_0$           & float       & Primary star mass [solar mass]         \\%\midrule
	  secondary mass $m_1$           & float       & Secondary star mass [solar mass]       \\%\midrule
	  bary pmra      $\dot{\alpha}$  & double      & Barycenter RA proper motion [mas/yr]   \\%\midrule
	  bary pmdec     $\dot{\delta}$  & double      & Barycenter DEC proper motion [mas/yr]  \\\bottomrule
	\end{tabularx}
\caption{Format of \file{binary\_orbitparam.dat}}
\label{tab:StarCatalogues:BinaryOrbitalParameters:file}
\end{table}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "guide"
%%% End: 
