%\chapterimage{chapter-t1-bg} % Chapter heading image

\chapter{Configuration File}
\label{ch:ConfigurationFile}
See \ref{ch:AdvancedUse:TheMainConfigurationFile}{Advanced Use, The
Main Configuration File} for information about this file, including its
default installed location, and command line options that can
affect how it is processed.

%Deprecated parameters are marked by gray background. 
%Possible new parameters are marked by yellow background.
%% GZ This is no longer true. Reactivate somehow? 

The file \file{config.ini} (or a file which you can load instead with the \file{-\/-config <file>} option) is structured into the following parts.

\section{\big[astro\big]}
\label{sec:config.ini:astro}

\begin{longtabu} to \textwidth {l|l|X}
\toprule
\emph{ID} & \emph{Type} & \emph{Description}\tabularnewline
\midrule
apparent\_magnitude\_algorithm & string & Set
algorithm for computation of apparent magnitude of the planets. Possible
values: \emph{Planesas}, \emph{Mueller}, \emph{Harris} and
\emph{Generic}\footnote{Available since version 0.13.3}. Default value: \emph{Harris}.\tabularnewline
\midrule
nebula\_magnitude\_limit & float & Value of limit
magnitude of the deep-sky objects. Default value: \emph{8.5}.\tabularnewline
star\_magnitude\_limit & float & Value of limit magnitude of
the stars\tabularnewline
\midrule
flag\_nebula\_magnitude\_limit & bool & Set to
\emph{true} to set limits for showing deep-sky objects. Default value: \emph{false}.\tabularnewline
\midrule
flag\_star\_magnitude\_limit & bool & Set to \emph{true} to
set limits for showing stars\tabularnewline
\midrule
flag\_extinction\_below\_horizon & bool & Set to \emph{true}
to apply extinction effects to sky below horizon\tabularnewline
\midrule
extinction\_mode\_below\_horizon & string & Set extinction
mode for atmosphere below horizon. Possible values: \emph{zero},
\emph{mirror} and \emph{max}. Default value is
\emph{zero}.\tabularnewline
\midrule
flag\_stars & bool & Set to \emph{false} to hide the
stars on start-up\tabularnewline
\midrule
flag\_star\_name & bool & Set to \emph{false} to hide the
star labels on start-up\tabularnewline
\midrule
flag\_planets & bool & Set to \emph{false} to hide the
planet labels on start-up\tabularnewline
\midrule
flag\_planets\_hints & bool & Set to \emph{false} to hide
the planet hints on start-up (names and circular
highlights)\tabularnewline
\midrule
flag\_planets\_orbits & bool & Set to \emph{true} to show
the planet orbits on start-up\tabularnewline
\midrule
flag\_permanent\_orbits & bool & Set to \emph{true} to show
the orbit of planet, when planet is out of the viewport
also.\tabularnewline
\midrule
flag\_planets\_pointers & bool & Set to \emph{true} to show
the planet pointer markers on start-up\tabularnewline
\midrule
flag\_light\_travel\_time & bool & Set to \emph{true} to
improve accuracy in the movement of the planets by compensating for the
time it takes for light to travel. This has an impact on
performance.\tabularnewline
\midrule
flag\_object\_trails & bool & Turns on and off drawing of
object trails (which show the movement of the planets over
time)\tabularnewline
\midrule
flag\_nebula & bool & Set to \emph{false} to
hide the nebulae on start-up. Default value: \emph{true}.\tabularnewline
\midrule
flag\_nebula\_name & bool & Set to
\emph{true} to show the nebula labels on start-up. Default value: \emph{false}.\tabularnewline
\midrule
flag\_nebula\_long\_name & bool & Set to \emph{true} to show
the nebula long labels on start-up.\tabularnewline
\midrule
flag\_nebula\_display\_no\_texture & bool & Set
to \emph{true} to suppress displaying of nebula textures. Default value: \emph{false}.\tabularnewline
\midrule
nebula\_hints\_amount & float & Sets the amount of
hints (between 0 and 10). Default value: \emph{3.0}.\tabularnewline
\midrule
nebula\_labels\_amount & float & Sets the amount of
labels (between 0 and 10). Default value: \emph{3.0}.\tabularnewline
\midrule
flag\_milky\_way & bool & Set to \emph{false}
to hide the Milky Way. Default value: \emph{true}.\tabularnewline
\midrule
milky\_way\_intensity & float & Sets the relative
brightness with which the milky way is drawn. Typical value: \emph{1} to
\emph{10}. Default value: \emph{1.0}.\tabularnewline
\midrule
flag\_zodiacal\_light & bool & Sets to \emph{false} to hide
the zodiacal light\tabularnewline
\midrule
zodiacal\_light\_intensity & float & Sets the relative
brightness with which the zodiacal light is drawn. Typical value:
\emph{1} to \emph{10}.\tabularnewline
\midrule
max\_mag\_nebula\_name & float & Sets the magnitude of
the nebulae whose name is shown. Typical value: \emph{8}\tabularnewline
\midrule
nebula\_scale & float & Sets how much to scale nebulae. a
setting of \emph{1} will display nebulae at normal size. Default value: \emph{1.0}.\tabularnewline
\midrule
flag\_bright\_nebulae & bool & Set to \emph{true} to increase
nebulae brightness to enhance viewing (less realistic)\tabularnewline
\midrule
flag\_nebula\_ngc & bool & Enables/disables display of all
NGC objects\tabularnewline
\midrule
flag\_nebula\_hints\_proportional & bool &
Enables/disables proportional markers for deep-sky
objects. Default value: \emph{false}.\tabularnewline
\midrule
flag\_surface\_brightness\_usage & bool &
Enables/disables usage surface brightness for markers for deep-sky
objects. Default value: \emph{false}.\tabularnewline
\midrule
flag\_use\_type\_filter & bool &
Enables/disables usage of the type filters for deep-sky
objects. Default value: \emph{false}.\tabularnewline
\bottomrule
\end{longtabu}

\section{\big[color\big]}
\label{sec:config.ini:color}

This section defines the RGB colors for the various objects, lines,
grids, labels etc. Values are given in float from 0 to
1. e.g. \emph{1.0,1.0,1.0} for white, or \emph{1,0,0} for red. Leave
no whitespace between the numbers!

\begin{longtabu} to \textwidth {l|l|X}
\toprule
\emph{ID} & \emph{Type} & \emph{Description}\tabularnewline
\midrule
default\_color & float R,G,B & default colour.  Default value: \emph{0.5,0.5,0.7}. \tabularnewline
\midrule
azimuthal\_color & float R,G,B & colour of the azimuthal grid. Default value: \emph{0.3,0.2,0.1}.\tabularnewline
\midrule
equatorial\_color & float R,G,B & colour of the equatorial grid (of date). Default value: \emph{0.2,0.3,0.8}.\tabularnewline
\midrule
equatorial\_J2000\_color & float R,G,B & colour of the equatorial grid (J2000). Default value: \emph{0.1,0.1,0.5}.\tabularnewline
\midrule
galactic\_color & float R,G,B & colour of the galactic grid. Default value: \emph{0.0,0.3,0.2}.\tabularnewline
\midrule
galactic\_equator\_color & float R,G,B & colour of the galactic equator line. Default value: \emph{0.5,0.3,0.1}. \tabularnewline
\midrule
equator\_color & float R,G,B & colour of the equatorial line. Default value: \emph{0.3,0.5,1.0}. \tabularnewline
\midrule
ecliptic\_color & float R,G,B & colour of the ecliptic line (J2000). Default value: \emph{0.9,0.6,0.2}. \tabularnewline
\midrule
ecliptic\_J2000\_color & float R,G,B & colour of the ecliptic grid (J2000). Default value: \emph{0.7,0.2,0.2}. \tabularnewline
\midrule
meridian\_color & float R,G,B & colour of the meridian line. Default value: \emph{0.2,0.6,0.2}. \tabularnewline
\midrule
horizon\_color & float R,G,B & colour of the horizon line. Default value: \emph{0.2,0.6,0.2}. \tabularnewline
\midrule
const\_lines\_color & float R,G,B & colour of the constellation lines. Default value: \emph{0.2,0.2,0.6}. \tabularnewline
\midrule
const\_names\_color & float R,G,B & colour of the constellation names. Default value: \emph{0.4,0.6,0.9}. \tabularnewline
\midrule
const\_boundary\_color & float R,G,B & colour of the constellation boundaries. Default value: \emph{0.3,0.1,0.1}. \tabularnewline
\midrule
star\_label\_color & float R,G,B & colour of the star labels. Default value: \emph{0.4,0.3,0.5}. \tabularnewline
\midrule
cardinal\_color & float R,G,B &  colour of the cardinal points. Default value: \emph{0.8,0.2,0.1}. \tabularnewline
\midrule
planet\_names\_color & float R,G,B & colour of the planet names. Default value: \emph{0.5,0.5,0.7}. \tabularnewline
\midrule
planet\_orbits\_color & float R,G,B & colour of the orbits. Default value: \emph{0.7,0.2,0.2}. \tabularnewline
\midrule
planet\_pointers\_color & float R,G,B & colour of the planet pointers. Default value: \emph{1.0,0.3,0.3}. \tabularnewline
\midrule
object\_trails\_color & float R,G,B & colour of the planet trails. Default value: \emph{1.0,0.7,0.0}. \tabularnewline
\midrule
telescope\_circle\_color & float R,G,B & colour of the telescope location indicator. Default value: \emph{0.6,0.4,0.0}. \tabularnewline
\midrule
telescope\_label\_color & float R,G,B & colour of the telescope location label. Default value: \emph{0.6,0.4,0.0}. \tabularnewline
\midrule
script\_console\_keyword\_color & float R,G,B & colour of the keywords in the script console. Default value: \emph{1.0,0.0,1.0}. \tabularnewline
\midrule
script\_console\_module\_color & float R,G,B & colour of the modules in the script console. Default value: \emph{0.0,1.0,1.0}. \tabularnewline
\midrule
script\_console\_comment\_color & float R,G,B & colour of the comments in the script console. Default value: \emph{1.0,1.0,0.0}. \tabularnewline
\midrule
script\_console\_function\_color & float R,G,B & colour of the functions in the script console. Default value: \emph{0.0,1.0,0.0}. \tabularnewline
\midrule
script\_console\_constant\_color & float R,G,B & colour of the constants in the script console. Default value: \emph{1.0,0.5,0.5}. \tabularnewline
\midrule
daylight\_text\_color & float R,G,B & colour of the info text at daylight. Default value: \emph{0.0,0.0,0.0}. \tabularnewline
\midrule
dso\_label\_color & float R,G,B & colour of the deep-sky objects labels. Default value: \emph{0.2,0.6,0.7}. \tabularnewline
\midrule
dso\_circle\_color & float R,G,B & colour of the deep-sky objects symbols. Default value: \emph{1.0,0.7,0.2}. \tabularnewline
\midrule
dso\_galaxy\_color & float R,G,B & colour of the galaxies symbols. Default value: \emph{1.0,0.2,0.2}. \tabularnewline
\midrule
dso\_radio\_galaxy\_color & float R,G,B & colour of the radio galaxies symbols. Default value: \emph{0.3,0.3,0.3}. \tabularnewline
\midrule
dso\_active\_galaxy\_color & float R,G,B & colour of the active galaxies symbols. Default value: \emph{1.0,0.5,0.2}. \tabularnewline
\midrule
dso\_interacting\_galaxy\_color & float R,G,B & colour of the interacting galaxies symbols. Default value: \emph{0.2,0.5,1.0}. \tabularnewline
\midrule
dso\_quasar\_color & float R,G,B & colour of the quasars symbols. Default value: \emph{1.0,0.2,0.2}. \tabularnewline
\midrule
dso\_possible\_quasar\_color & float R,G,B & colour of the possible quasars symbols. Default value: \emph{1.0,0.2,0.2}. \tabularnewline
\midrule
dso\_bl\_lac\_color & float R,G,B & colour of the BL Lac objects symbols. Default value: \emph{1.0,0.2,0.2}. \tabularnewline
\midrule
dso\_blazar\_color & float R,G,B & colour of the blazars symbols. Default value: \emph{1.0,0.2,0.2}. \tabularnewline
\midrule
dso\_nebula\_color & float R,G,B & colour of the nebulae symbols. Default value: \emph{0.1,1.0,0.1}. \tabularnewline
\midrule
dso\_planetary\_nebula\_color & float R,G,B & colour of the planetary nebulae symbols. Default value: \emph{0.1,1.0,0.1}. \tabularnewline
\midrule
dso\_reflection\_nebula\_color & float R,G,B & colour of the reflection nebulae symbols. Default value: \emph{0.1,1.0,0.1}. \tabularnewline
\midrule
dso\_bipolar\_nebula\_color & float R,G,B & colour of the bipolar nebulae symbols. Default value: \emph{0.1,1.0,0.1}. \tabularnewline
\midrule
dso\_emission\_nebula\_color & float R,G,B & colour of the emission nebulae symbols. Default value: \emph{0.1,1.0,0.1}. \tabularnewline
\midrule
dso\_dark\_nebula\_color & float R,G,B & colour of the dark nebulae symbols. Default value: \emph{0.3,0.3,0.3}. \tabularnewline
\midrule
dso\_hydrogen\_region\_color & float R,G,B & colour of the hydrogen regions symbols. Default value: \emph{0.1,1.0,0.1}. \tabularnewline
\midrule
dso\_supernova\_remnant\_color & float R,G,B & colour of the supernovae remnants symbols. Default value: \emph{0.1,1.0,0.1}. \tabularnewline
\midrule
dso\_interstellar\_matter\_color & float R,G,B & colour of the interstellar matter symbols. Default value: \emph{0.1,1.0,0.1}. \tabularnewline
\midrule
dso\_cluster\_with\_nebulosity\_color & float R,G,B & colour of the clusters associated with nebulosity symbols. Default value: \emph{0.1,1.0,0.1}. \tabularnewline
\midrule
dso\_molecular\_cloud\_color & float R,G,B & colour of the molecular clouds symbols. Default value: \emph{0.1,1.0,0.1}. \tabularnewline
\midrule
dso\_possible\_planetary\_nebula\_color & float R,G,B & colour of the possible planetary nebulae symbols. Default value: \emph{0.1,1.0,0.1 }. \tabularnewline
\midrule
dso\_protoplanetary\_nebula\_color & float R,G,B & colour of the protoplanetary nebulae symbols. Default value: \emph{0.1,1.0,0.1}. \tabularnewline
\midrule
dso\_cluster\_color & float R,G,B & colour of the star clusters symbols. Default value: \emph{1.0,1.0,0.1}. \tabularnewline
\midrule
dso\_open\_cluster\_color & float R,G,B & colour of the open star clusters symbols. Default value: \emph{1.0,1.0,0.1}. \tabularnewline
\midrule
dso\_globular\_cluster\_color & float R,G,B & colour of the globular star clusters symbols. Default value: \emph{1.0,1.0,0.1}. \tabularnewline
\midrule
dso\_stellar\_association\_color & float R,G,B & colour of the stellar associations symbols. Default value: \emph{1.0,1.0,0.1}. \tabularnewline
\midrule
dso\_star\_cloud\_color & float R,G,B & colour of the star clouds symbols. Default value: \emph{1.0,1.0,0.1}. \tabularnewline
\midrule
dso\_star\_color & float R,G,B & colour of the star symbols. Default value: \emph{1.0,0.7,0.2}. \tabularnewline
\midrule
dso\_emission\_object\_color & float R,G,B & colour of the emission objects symbols. Default value: \emph{1.0,0.7,0.2}. \tabularnewline
\midrule
dso\_young\_stellar\_object\_color & float R,G,B & colour of the young stellar objects symbols. Default value: \emph{1.0,0.7,0.2}. \tabularnewline
\bottomrule
\end{longtabu}

\section{\big[custom\_selected\_info\big]}
\label{sec:config.ini:custom_selected_info}

You can fine-tune the bits of information to display for the selected object in this section. Set the entry to \emph{true} to display it.

\begin{longtabu} to \textwidth {l|l|X}
\toprule
\emph{ID} & \emph{Type} & \emph{Description}\tabularnewline
\midrule
flag\_show\_absolutemagnitude & bool & absolute magnitude for objects.\tabularnewline
\midrule
flag\_show\_altaz & bool & horizontal coordinates for objects.\tabularnewline
\midrule
flag\_show\_catalognumber & bool & catalog designations for objects.\tabularnewline
\midrule
flag\_show\_distance & bool & distance to object.\tabularnewline
\midrule
flag\_show\_extra & bool & extra info for object.\tabularnewline
\midrule
flag\_show\_hourangle & bool & hour angle for object.\tabularnewline
\midrule
flag\_show\_magnitude & bool & magnitude for object.\tabularnewline
\midrule
flag\_show\_name & bool & common name for object.\tabularnewline
\midrule
flag\_show\_radecj2000 & bool & equatorial coordinates (J2000) of object.\tabularnewline
\midrule
flag\_show\_radecofdate & bool & equatorial coordinates (of date) of object.\tabularnewline
\midrule
flag\_show\_size & bool & size of object.\tabularnewline
\midrule
flag\_show\_galcoord & bool & galactic coordinates (System~II) of object.\tabularnewline
\midrule
flag\_show\_eclcoord & bool & ecliptic coordinates (J2000 and of date) of object.\tabularnewline
\midrule
flag\_show\_type & bool & type of object\tabularnewline
\bottomrule
\end{longtabu}

%% TODO: Maybe separate flag_show_eclcoord and flag_show_eclcoordj2000?

\section{\big[custom\_time\_correction\big]}
\label{sec:config.ini:custom_time_correction}

\begin{longtabu} to \textwidth {l|l|X}
\toprule
\emph{ID} & \emph{Type} & \emph{Description}\tabularnewline
\midrule
coefficients & {[}float,float,float{]} & Coefficients for
custom equation of DeltaT\tabularnewline
\midrule
ndot & float & n-dot value for custom equation of
DeltaT\tabularnewline
\midrule
year & int & Year for custom equation of DeltaT\tabularnewline
\bottomrule
\end{longtabu}

\section{\big[devel\big]}
\label{sec:config.ini:devel}

\begin{longtabu} to \textwidth {l|l|X}
\toprule
\emph{ID} & \emph{Type} & \emph{Description}\tabularnewline
\midrule
convert\_dso\_catalog & bool & Set to
\emph{true} to convert \texttt{catalog.txt} file into
\texttt{catalog.dat} file. Default value: \emph{false}.\tabularnewline
\midrule
convert\_dso\_decimal\_coord & bool & Set to
\emph{true} to use decimal values for coordinates in source
catalog. Default value: \emph{true}.\tabularnewline
\bottomrule
\end{longtabu}

\section{\big[dso\_catalog\_filters\big]}
\label{sec:config.ini:dso_catalog_filters}
In this section you can fine-tune which of the deep-sky catalogs should be activated on startup.

\begin{longtabu} to \textwidth {l|l|X}
\toprule
\emph{ID} & \emph{Type} & \emph{Description}\tabularnewline
\midrule
flag\_show\_ngc & bool & New General Catalogue (NGC). Default value: \emph{true}. \tabularnewline
\midrule
flag\_show\_ic  & bool & Index Catalogue (IC). Default value: \emph{true}. \tabularnewline
\midrule
flag\_show\_m   & bool & Messier Catalog (M). Default value: \emph{true}. \tabularnewline
\midrule
flag\_show\_c   & bool & Caldwell Catalogue (C). Default value: \emph{false}. \tabularnewline
\midrule
flag\_show\_b   & bool & Barnard Catalogue (B). Default value: \emph{false}. \tabularnewline
\midrule
flag\_show\_sh2 & bool & Sharpless Catalogue (Sh2). Default value: \emph{false}. \tabularnewline
\midrule
flag\_show\_vdb & bool & Van den Bergh Catalogue of reflection nebulae (VdB). Default value: \emph{false}. \tabularnewline
\midrule
flag\_show\_rcw & bool & The RCW catalogue of H$\alpha$-emission regions in the southern Milky Way. Default value: \emph{false}. \tabularnewline
\midrule
flag\_show\_lbn & bool & Lynds' Catalogue of Bright Nebulae (LBN). Default value: \emph{false}. \tabularnewline
\midrule
flag\_show\_ldn & bool & Lynds' Catalogue of Dark Nebulae (LDN). Default value: \emph{false}. \tabularnewline
\midrule
flag\_show\_cr  & bool & Collinder Catalogue (Cr). Default value: \emph{false}. \tabularnewline
\midrule
flag\_show\_mel & bool & Melotte Catalogue of Deep Sky Objects (Mel). Default value: \emph{false}. \tabularnewline
\midrule
flag\_show\_pgc & bool & HYPERLEDA. I. Catalog of galaxies (PGC). Default value: \emph{false}. \tabularnewline
\midrule
flag\_show\_ced & bool & Cederblad Catalog of bright diffuse Galactic nebulae (Ced). Default value: \emph{false}. \tabularnewline
\midrule
flag\_show\_ugc & bool & The Uppsala General Catalogue of Galaxies (UGC). Default value: \emph{false}. \tabularnewline
\bottomrule
\end{longtabu}

\section{\big[dso\_type\_filters\big]}
\label{section-dsoux5ftypeux5ffilters}

\begin{longtabu} to \textwidth {l|l|X}
\toprule
\emph{ID} & \emph{Type} & \emph{Description}\tabularnewline
\midrule
flag\_show\_galaxies & bool & Set to \emph{true}
to displaying galaxies. Default value: \emph{true}. \tabularnewline
\midrule
flag\_show\_active\_galaxies & bool & Set to
\emph{true} to displaying active galaxies. Default value: \emph{true}. \tabularnewline
\midrule
flag\_show\_interacting\_galaxies & bool & Set
to \emph{true} to displaying interacting galaxies. Default value: \emph{true}. \tabularnewline
\midrule
flag\_show\_clusters & bool & Set to \emph{true}
to displaying star clusters. Default value: \emph{true}. \tabularnewline
\midrule
flag\_show\_bright\_nebulae & bool & Set to
\emph{true} to displaying bright nebulae. Default value: \emph{true}. \tabularnewline
\midrule
flag\_show\_dark\_nebulae & bool & Set to
\emph{true} to displaying dark nebulae. Default value: \emph{true}. \tabularnewline
\midrule
flag\_show\_planetary\_nebulae & bool & Set to
\emph{true} to displaying planetary nebulae. Default value: \emph{true}. \tabularnewline
\midrule
flag\_show\_hydrogen\_regions & bool & Set to
\emph{true} to displaying hydrogen regions. Default value: \emph{true}. \tabularnewline
\midrule
flag\_show\_supernova\_remnants & bool & Set to
\emph{true} to displaying supernovae remnants. Default value: \emph{true}. \tabularnewline
\midrule
flag\_show\_other & bool & Set to \emph{true} to
displaying other deep-sky objects. Default value: \emph{true}. \tabularnewline
\bottomrule
\end{longtabu}

\section{\big[gui\big]}\label{section-gui}

\begin{longtabu} to \textwidth {l|l|X}
\toprule
\emph{ID} & \emph{Type} & \emph{Description}\tabularnewline
\midrule
base\_font\_size & int & Sets the font size. Typical value:
\emph{15}\tabularnewline
\midrule
base\_font\_name & string & Selects the name for base font, e.g. \emph{DejaVu
Sans}\tabularnewline
\midrule
safe\_font\_name & string & Selects the name for safe font,
e.g. \emph{Verdana}\tabularnewline
\midrule
base\_font\_file & string & Selects the name for font file,
e.g. \emph{DejaVuSans.ttf}\tabularnewline
\midrule
flag\_show\_fps & bool & Set to \emph{false} if you don't
want to see at how many frames per second Stellarium is
rendering\tabularnewline
\midrule
flag\_show\_fov & bool & Set to \emph{false} if you don't
want to see how many degrees your field of view is\tabularnewline
\midrule
flag\_mouse\_cursor\_timeout & float & Set to \emph{0} if you
want to keep the mouse cursor visible at all times. non-0 values mean
the cursor will be hidden after that many seconds of
inactivity\tabularnewline
\midrule
flag\_show\_flip\_buttons & bool & Enables/disables display of
the image flipping buttons in the main toolbar (see section
\ref{imageflipping})\tabularnewline
\midrule
flag\_show\_nebulae\_background\_button & bool & Set to
\emph{true} if you want to have access to the button for
enabling/disabling backgrounds for deep-sky objects\tabularnewline
\midrule
flag\_use\_degrees & bool &\tabularnewline
\midrule
selected\_object\_info & string & Values: \emph{all},
\emph{short} and \emph{none}. Value \emph{custom} added since version
0.12.0.\tabularnewline
\midrule
auto\_hide\_horizontal\_toolbar & bool & Set to \emph{true} if
you want auto hide the horizontal toolbar.\tabularnewline
\midrule
auto\_hide\_vertical\_toolbar & bool & Set to \emph{true} if
you want auto hide the vertical toolbar.\tabularnewline
\midrule
day\_key\_mode & string & Specifies the amount of time which is
added and subtracted when the {[} {]} - and = keys are pressed -
calendar days, or sidereal days. This option only makes sense for
Digitalis planetariums. Values: \emph{calendar} or
\emph{sidereal}\tabularnewline
\midrule
flag\_use\_window\_transparency & bool & Set to \emph{false}
if you want see the opacity for bars\tabularnewline
\midrule
flag\_show\_datetime & bool & Set to \emph{true} if you want
display date and time in the bottom bar\tabularnewline
\midrule
flag\_time\_jd & bool & Set to \emph{true} if you want using
JD format for time in the bottom bar\tabularnewline
\midrule
flag\_show\_location & bool & Set to \emph{true} if you want
display location in the bottom bar\tabularnewline
\midrule
flag\_fov\_dms & bool & Set to \emph{true} if you want using
DMS format for FOV in the bottom bar\tabularnewline
\midrule
flag\_show\_decimal\_degrees & bool & Set to \emph{true} if
you want use decimal degrees for coordinates\tabularnewline
\midrule
flag\_use\_azimuth\_from\_south & bool & Set to \emph{true} if
you want calculate azimuth from south towards west (as in old
astronomical literature)\tabularnewline
\midrule
flag\_show\_gui & bool & Set to \emph{true} if you want
display GUI\tabularnewline
\bottomrule
\end{longtabu}

\section{\big[init\_location\big]}\label{section-init-location}

\begin{longtabu} to \textwidth {l|l|X}
\toprule
\emph{ID} & \emph{Type} & \emph{Description}\tabularnewline
\midrule
landscape\_name & string & Sets the landscape you see.
Other options are \emph{garching, guereins, trees, moon, ocean,
hurricane, hogerielen}\tabularnewline
\midrule
location & string & Name of location on which to start
stellarium.\tabularnewline
\midrule
last\_location & string & Coordinates of last used location in
stellarium.\tabularnewline
\bottomrule
\end{longtabu}

\section{\big[landscape\big]}\label{section-landscape}

\begin{longtabu} to \textwidth {l|l|X}
\toprule
\emph{ID} & \emph{Type} & \emph{Description}\tabularnewline
flag\_landscape & bool & Set to false if you don't want to
see the landscape at all\tabularnewline
\midrule
flag\_fog & bool & Set to \emph{false} if you don't want to
see fog on start-up\tabularnewline
\midrule
flag\_atmosphere & bool & Set to \emph{false} if you don't
want to see atmosphere on start-up\tabularnewline
\midrule
flag\_landscape\_sets\_location & bool & Set to \emph{true}
if you want Stellarium to modify the observer location when a new
landscape is selected (changes planet and longitude/latitude/altitude if
that data is available in the landscape.ini file)\tabularnewline
\midrule
minimal\_brightness & float & Set minimal brightness for
landscapes. Typical value: \emph{0.01}\tabularnewline
\midrule
atmospheric\_extinction\_coefficient & float & Set atmospheric
extinction coefficient\tabularnewline
\midrule
temperature\_C & float & Set atmospheric temperature
(Celsius)\tabularnewline
\midrule
pressure\_mbar & float & Set atmospheric pressure
(mbar)\tabularnewline
\bottomrule
\end{longtabu}

\section{\big[localization\big]}\label{section-localization}

\begin{longtabu} to \textwidth {l|l|X}
\toprule
\emph{ID} & \emph{Type} & \emph{Description}\tabularnewline
\midrule
sky\_culture & string & Sets the sky culture to use. E.g.
\emph{western, polynesian, egyptian, chinese, lakota, navajo, inuit,
korean, norse, tupi, maori, aztec, sami}.\tabularnewline
\midrule
sky\_locale & string & Sets language used for names of
objects in the sky (e.g. planets). The value is a short locale code,
e.g. \emph{en, de, en\_GB}\tabularnewline
\midrule
app\_locale & string & Sets language used for Stellarium's
user interface. The value is a short locale code, e.g. \emph{en, de,
en\_GB}.\tabularnewline
\midrule
time\_zone & string & Sets the time zone. Valid values:
\emph{system\_default}, or some region/location combination, e.g.
\emph{Pacific/Marquesas}\tabularnewline
\midrule
time\_display\_format & string & Set the time display format
mode: can be \emph{system\_default}, \emph{24h} or
\emph{12h}.\tabularnewline
\midrule
date\_display\_format & string & Set the date display format
mode: can be \emph{system\_default}, \emph{mmddyyyy}, \emph{ddmmyyyy} or
\emph{yyyymmdd} (ISO8601).\tabularnewline
\bottomrule
\end{longtabu}

\section{\big[main\big]}\label{section-main}

\begin{longtabu} to \textwidth {l|l|X}
\toprule
\emph{ID} & \emph{Type} & \emph{Description}\tabularnewline
\midrule
invert\_screenshots\_colors & bool & If \emph{true},
Stellarium will saving the screenshorts with inverted
colors.\tabularnewline
\midrule
restore\_defaults & bool & If \emph{true}, Stellarium will be
restore default settings at startup.\tabularnewline
\midrule
screenshot\_dir & string & Path for saving
screenshots\tabularnewline
\midrule
version & string & Version of Stellarium. This parameter
using for updating config.ini file\tabularnewline
\midrule
use\_separate\_output\_file & bool & Set to \emph{true} if you
want to create a new file for output for each start of
Stellarium\tabularnewline
\midrule
check\_requirements & bool & Set to \emph{false} if you want
to disable checking hardware requirements at startup. This options for
developers only!\tabularnewline
\bottomrule
\end{longtabu}

\section{\big[navigation\big]}\label{section-navigation}

\begin{longtabu} to \textwidth {l|l|X}
\toprule
\emph{ID} & \emph{Type} & \emph{Description}\tabularnewline
\midrule
preset\_sky\_time & float & Preset sky time used by the
dome version. Unit is Julian Day. Typical value:
\emph{2451514.250011573}\tabularnewline
\midrule
startup\_time\_mode & string & Set the start-up time mode,
can be \emph{actual} (start with current real world time), or
\emph{Preset} (start at time defined by
preset\_sky\_time)\tabularnewline
\midrule
flag\_enable\_zoom\_keys & bool & Set to \emph{false} if
you want to disable the zoom\tabularnewline
\midrule
flag\_manual\_zoom & bool & Set to'' false'' for normal
zoom behaviour as described in this guide. When set to true, the auto
zoom feature only moves in a small amount and must be pressed many
times\tabularnewline
\midrule
flag\_enable\_move\_keys & bool & Set to \emph{false} if
you want to disable the arrow keys\tabularnewline
\midrule
flag\_enable\_mouse\_navigation & bool & Set to
\emph{false} if you want to disable the mouse navigation.\tabularnewline
\midrule
init\_fov & float & Initial field of view, in degrees,
typical value:'' 60''\tabularnewline
\midrule
init\_view\_pos & floats & Initial viewing direction. This
is a vector with x,y,z-coordinates. x being N-S (S +ve), y being E-W (E
+ve), z being up-down (up +ve). Thus to look South at the horizon use
\emph{1,0,0}. To look Northwest and up at $45\degree$, use \emph{-1,-1,1} and so
on.\tabularnewline
\midrule
auto\_move\_duration & float & Duration for the program to
move to point at an object when the space bar is pressed. Typical value:
\emph{1.4}\tabularnewline
\midrule
mouse\_zoom & float & Sets the mouse zoom amount
(mouse-wheel)\tabularnewline
\midrule
move\_speed & float & Sets the speed of
movement\tabularnewline
\midrule
zoom\_speed & float & Sets the zoom speed\tabularnewline
\midrule
viewing\_mode & string & If set to \emph{horizon}, the
viewing mode simulate an alt/azi mount, if set to \emph{equator}, the
viewing mode simulates an equatorial mount\tabularnewline
\midrule
flag\_manual\_zoom & bool & Set to \emph{true} if you want
to auto-zoom in incrementally.\tabularnewline
\midrule
auto\_zoom\_out\_resets\_direction & bool & Set to
\emph{true} if you want to auto-zoom restoring direction.\tabularnewline
\midrule
time\_correction\_algorithm & string & Algorithm of DeltaT
correction.\tabularnewline
\bottomrule
\end{longtabu}

\section{\big[plugins\_load\_at\_startup\big]}
\label{sec:Reference:section-plugins-load-at-startup}

This section lists which plugins are loaded at startup (those with
\emph{true} values). Each plugin can add another section with its own
content, which is described in the respective plugin documentation,
see \ref{ch:Plugins}. You activate loading of plugins in the \key{F2}
settings dialog, tab ``Plugins''. After selection of which plugins to
load, you must restart Stellarium.

\begin{longtabu} to \textwidth {l|l|X}
\toprule
\emph{ID} & \emph{Type} & \emph{Description}\\
\midrule
AngleMeasure & bool & Angle Measure plugin\\
\midrule
CompassMarks & bool & Compass Marks plugin\\
\midrule
Exoplanets & bool & Exoplanets plugin \\
\midrule
Observability & bool & Observability Analysis\\
\midrule
Oculars & bool & Oculars plugin \\
\midrule
Pulsars & bool & Pulsars plugin \\
\midrule
Quasars & bool & Quasars plugin \\
\midrule
Satellites & bool & Satellites plugin \\
\midrule
SolarSystemEditor & bool & Solar System Editor plugin\\
\midrule
Supernovae & bool & Historical Supernovae plugin \\
\midrule
TelescopeControl & bool & Telescope Control plugin \\
\midrule
TextUserInterface & bool & Text User Interface plugin \\
\midrule
TimeZoneConfiguration & bool & Time Zone plugin \\
\midrule
Novae & bool & Bright Novae plugin \\
\midrule
Scenery3dMgr & bool & Scenery 3D plugin \\
\bottomrule
\end{longtabu}

\section{\big[projection\big]}\label{section-projection}

This section contains the projection of your choice and several
advanced settings important if you run Stellarium on a single screen,
multi-projection, dome projection, or other setups.

\begin{longtabu} to \textwidth {l|l|X}
\toprule
\emph{ID} & \emph{Type} & \emph{Description}\tabularnewline
\midrule
type & string & Sets projection mode. Values: \emph{ProjectionPerspective,
ProjectionEqualArea, ProjectionStereographic, ProjectionFisheye,
ProjectionHammer, ProjectionCylinder, ProjectionMercator},
\emph{ProjectionOrthographic}, or \emph{StelProjectorSinusoidal}.\tabularnewline
\midrule
viewport & string & How the view-port looks. Values:
\emph{none} (regular rectangular screen), \emph{disk} (circular screen, useful for planetarium setup).\tabularnewline
\midrule
viewportMask & string & How the view-port looks. Values: \emph{none}.\tabularnewline
\midrule
flag\_use\_gl\_point\_sprite & bool & (deprecated.)\  \tabularnewline
\midrule
flip\_horz & bool & \tabularnewline
\midrule
flip\_vert & bool & \tabularnewline
\midrule
viewport\_center\_offset\_x & float & [-0.5\ldots+0.5] Usually 0. \\
\midrule
viewport\_center\_offset\_y & float & [-0.5\ldots+0.5] Use negative values to move the horizon lower. \\
\bottomrule
\end{longtabu}

\section{\big[proxy\big]}\label{section-proxy}

\begin{longtabu} to \textwidth {l|l|X}
\toprule
\emph{ID} & \emph{Type} & \emph{Description}\tabularnewline
\midrule
host\_name & string & Name of host for proxy. E.g. \emph{proxy.org}\tabularnewline
\midrule
port & int & Port of proxy. E.g. \emph{8080}\tabularnewline
\midrule
user & string & Username for proxy. E.g. \emph{michael\_knight}\tabularnewline
\midrule
password & string & Password for proxy. E.g. \emph{xxxxx}\tabularnewline
\bottomrule
\end{longtabu}

\section{\big[scripts\big]}\label{section-scripts}

\begin{longtabu} to \textwidth {l|l|X}
\toprule
\emph{ID} & \emph{Type} & \emph{Description}\tabularnewline
\midrule
flag\_script\_allow\_ui & bool &\tabularnewline
\midrule
scripting\_allow\_write\_files & bool &\tabularnewline
\bottomrule
\end{longtabu}

\section{\big[search\big]}\label{section-search}

\begin{longtabu} to \textwidth {l|l|X}
\toprule
\emph{ID} & \emph{Type} & \emph{Description}\tabularnewline
\midrule
flag\_search\_online & bool & If \emph{true}, Stellarium will be use SIMBAD for search.\tabularnewline
\midrule
simbad\_server\_url & string & URL for SIMBAD mirror\tabularnewline
\midrule
flag\_start\_words & bool & If \emph{false}, Stellarium will be search phrase only from start of words\tabularnewline
\midrule
coordinate\_system & string & Specifies the coordinate system. \emph{Possible values:} equatorialJ2000, equatorial, horizontal,
galactic. \emph{Default value:} equatorialJ2000.\tabularnewline
\bottomrule
\end{longtabu}

\section{\big[spheric\_mirror\big]}\label{section-sphericux5fmirror}

\begin{longtabu} to \textwidth {l|l|X}
\toprule
\emph{ID} & \emph{Type} & \emph{Description}\tabularnewline
\midrule
distorter\_max\_fov & float & Set the maximum field of view
for the spheric mirror distorter in degrees. Typical value,
\emph{180}\tabularnewline
\midrule
flag\_use\_ext\_framebuffer\_object & bool & Some video
hardware incorrectly claims to support some GL extension,
GL\_FRAMEBUFFER\_EXTEXT. If, when using the spheric mirror distorter the
frame rate drops to a very low value (e.g. 0.1 FPS), set this parameter
to false to tell Stellarium ignore the claim of the video driver that it
can use this extension\tabularnewline
\midrule
flip\_horz & bool & Flip the projection horizontally\tabularnewline
\midrule
flip\_vert & bool & Flip the projection vertically\tabularnewline
\midrule
projector\_gamma & float & This parameter controls the
properties of the spheric mirror projection mode.\tabularnewline
\midrule
projector\_position\_x & float & \tabularnewline
\midrule
projector\_position\_y & float & \tabularnewline
\midrule
projector\_position\_z & float & \tabularnewline
\midrule
mirror\_position\_x & float & \tabularnewline
\midrule
mirror\_position\_y & float & \tabularnewline
\midrule
mirror\_position\_z & float & \tabularnewline
\midrule
mirror\_radius & float & \tabularnewline
\midrule
dome\_radius & float & \tabularnewline
\midrule
zenith\_y & float & \tabularnewline
\midrule
scaling\_factor & float & \tabularnewline
\bottomrule
\end{longtabu}

\section{\big[stars\big]}\label{section-stars}

\begin{longtabu} to \textwidth {l|l|X}
\toprule
\emph{ID} & \emph{Type} & \emph{Description}\tabularnewline
\midrule
relative\_scale & float & Changes the relative size of
bright and faint stars. Higher values mean that bright stars are
comparitively larger when rendered. Typical value:
\emph{1.0}\tabularnewline
\midrule
absolute\_scale & float & Changes how large stars are
rendered. larger value lead to larger depiction. Typical value:
\emph{1.0}\tabularnewline
\midrule
star\_twinkle\_amount & float & Sets the amount of
twinkling. Typical value: \emph{0.3}\tabularnewline
\midrule
flag\_star\_twinkle & bool & Set to \emph{false} to turn
star twinkling off, \emph{true} to allow twinkling.\tabularnewline
\midrule
mag\_converter\_max\_fov & float & Sets the maximum field
of view for which the magnitude conversion routine is used. Typical
value: \emph{90.0}.\tabularnewline
\midrule
mag\_converter\_min\_fov & float & Sets the minimum field
of view for which the magnitude conversion routine is used. Typical
value: \emph{0.001}.\tabularnewline
\midrule
labels\_amount & float & Sets the amount of labels. Typical
value: \emph{3.0}\tabularnewline
\midrule
init\_bortle\_scale & int & Sets the initial value of the
bortle scale. Typical value: \emph{3}.\tabularnewline
\bottomrule
\end{longtabu}

\section{\big[tui\big]}\label{section-tui}

\begin{longtabu} to \textwidth {l|l|X}
\toprule
\emph{ID} & \emph{Type} & \emph{Description}\tabularnewline
\midrule
flag\_enable\_tui\_menu & bool & Enables or disables the TUI menu\tabularnewline
\midrule
flag\_show\_gravity\_ui & bool & Enables or disables gravity mode for UI\tabularnewline
\midrule
flag\_show\_tui\_datetime & bool & Set to \emph{true} if you want to see a date and time label suited for dome projections\tabularnewline
\midrule
flag\_show\_tui\_short\_obj\_info & bool & set to \emph{true} if you want to see object info suited for dome projections\tabularnewline
\bottomrule
\end{longtabu}

\section{\big[video\big]}\label{section-video}

\begin{longtabu} to \textwidth {l|l|X}
\toprule
\emph{ID} & \emph{Type} & \emph{Description}\tabularnewline
\midrule
fullscreen & bool & If \emph{true}, Stellarium will start
up in full-screen mode. If \emph{false}, Stellarium will start in
windowed mode\tabularnewline
\midrule
screen\_w & int & Sets the display width when in windowed mode. Value in pixels, e.g. \emph{1024}\tabularnewline
\midrule
screen\_h & int & Sets the display height when in windowed mode. Value in pixels, e.g. \emph{768}\tabularnewline
\midrule
screen\_x & int & Value in pixels, e.g. \emph{0}\tabularnewline
\midrule
screen\_y & int & Value in pixels, e.g. \emph{0}\tabularnewline
\midrule
viewport\_effect & string & This is used when the spheric mirror display mode is activated. Values include \emph{none} and \emph{sphericMirrorDistorter}.\tabularnewline
\midrule
minimum\_fps & int & Sets the minimum number of frames per second to display at (hardware performance permitting)\tabularnewline
\midrule
maximum\_fps & int & Sets the maximum number of frames per second to display at. This is useful to reduce power consumption in laptops.\tabularnewline
\bottomrule
\end{longtabu}

\section{\big[viewing\big]}\label{section-viewing}
This section defines which objects, labels, lines, grids etc.\ you want to see on startup. Set those to \emph{true}.

\begin{longtabu} to \textwidth {l|l|X}
\toprule
\emph{ID} & \emph{Type} & \emph{Description}\tabularnewline
\midrule
atmosphere\_fade\_duration & float & Sets the time (seconds) it takes for the atmosphere to fade when de-selected\tabularnewline
\midrule
flag\_constellation\_drawing & bool & constellation line drawing\tabularnewline
\midrule
flag\_constellation\_name & bool & constellation names\tabularnewline
\midrule
flag\_constellation\_art & bool & constellation art\tabularnewline
\midrule
flag\_constellation\_boundaries & bool & constellation boundaries \tabularnewline
\midrule
flag\_constellation\_isolate\_selected & bool & When \emph{true}, constellation lines, boundaries and art will be limited to the constellation of the selected star, if that star is ``on'' one of the constellation lines.\tabularnewline
\midrule
flag\_constellation\_pick & bool & Set to \emph{true} if you only want to see the line drawing, art and name of the selected constellation star\tabularnewline
\midrule
flag\_isolated\_trails & bool & Set to \emph{true} if you only want to see the tail line drawing of the selected planets (asteroids, comets, moons)\tabularnewline
\midrule
flag\_azimutal\_grid & bool & azimuthal grid \tabularnewline
\midrule
flag\_equatorial\_grid & bool & equatorial grid (of date) \tabularnewline
\midrule
flag\_equatorial\_J2000\_grid & bool & equatorial grid (J2000) \tabularnewline
\midrule
flag\_ecliptic\_J2000\_grid & bool & ecliptic grid (J2000) \tabularnewline
\midrule
flag\_galactic\_grid & bool & galactic grid (System~II)\tabularnewline
\midrule
flag\_galactic\_equator\_line & bool & galactic equator line \tabularnewline
\midrule
flag\_equator\_J2000\_line & bool & equator line (J2000) \tabularnewline
\midrule
flag\_equator\_line & bool & equator line \tabularnewline
\midrule
flag\_ecliptic\_line & bool & ecliptic line (of date) \tabularnewline
\midrule
flag\_ecliptic\_J2000\_line & bool & ecliptic line (J2000) \tabularnewline
\midrule
flag\_meridian\_line & bool & meridian line \tabularnewline
\midrule
flag\_cardinal\_points & bool & cardinal points\tabularnewline
\midrule
flag\_gravity\_labels & bool & Set to \emph{true} if you want labels to undergo gravity (top side of text points toward zenith). Useful with dome projection.\tabularnewline
\midrule
flag\_moon\_scaled & bool & Change to \emph{false} if you want to see the real moon size \tabularnewline
\midrule
moon\_scale & float & Sets the moon scale factor, to correlate to our perception of the moon's size. Typical value: \emph{4}\tabularnewline
\midrule
constellation\_art\_intensity & float & brightness [0\ldots1] of the constellation art images. Typical value: \emph{0.5}\tabularnewline
\midrule
constellation\_art\_fade\_duration & float & time the constellation art takes to fade in or out, in seconds. Typical value: \emph{1.5}\tabularnewline
\midrule
constellation\_font\_size & int & font size for constellation labels\tabularnewline
\midrule
constellation\_line\_thickness & float & thickness of lines of the constellations [1\ldots5]. Typical value: \emph{1}\tabularnewline
\midrule
flag\_night & bool & Enable night mode (red-only mode) on startup\tabularnewline
\midrule
light\_pollution\_luminance & float & Sets the level of the light pollution simulation\tabularnewline %% TODO: WHAT IS THIS?
\midrule
use\_luminance\_adaptation & bool & Enable dynamic eye adaptation.\tabularnewline %% TODO: DESCRIBE!
\midrule
flag\_isolated\_trails & float & Change to \emph{false} if you want to see the trails for all celestial bodies from Solar system.\tabularnewline
\midrule
flag\_isolated\_orbits & float & Change to \emph{false} if you want to see orbits for selected planet and their moons.\tabularnewline
\bottomrule
\end{longtabu}





%%% Local Variables: 
%%% mode: latex
%%% TeX-PDF-mode: t
%%% TeX-master: "guide"
%%% End: 

