%% Part of Stellarium User Guide 0.15+
%% History:
%% 2016-04-12 adapted from Scenery3D plugin documentation. Just reformatted.
%% Doc should be edited in here from now on. 
%% !!!!!!!!!! Please ask GZ before editing this file. !!!!!!!!!!!!!!!

% \documentclass[a4paper]{article}
% \usepackage[latin1]{inputenc}
% \usepackage[T1]{fontenc}
% \usepackage{a4wide}
% \usepackage{hyperref}

%\newcommand{\filename}[1]{\texttt{#1}}
%\newcommand{\cmd}[1]{\texttt{#1}}

%\chapter{Scenery3d -- Walkable 3D Models in Stellarium}  % TOO LONG
\chapter{Scenery3d -- 3D Landscapes}% in Stellarium} 
\label{ch:scenery3d}
\chapterauthor*{Georg Zotti and Florian Schaukowitsch}
%\footnote{\url{Georg.Zotti@univie.ac.at},
%}\  \

%\date{April 10, 2015; last updated 20160412}
%\begin{document}
%\maketitle

\section{Introduction}
\label{sec:scenery3d:Introduction}


Have you ever wished to be able to walk through Stonehenge or other
ancient building structures described as being constructed with
astronomical orientation in mind, and experience such orientation in a
3D virtual environment that also provides a good sky simulation?

The Stellarium Scenery3d plugin allows you to see architectural 3D models
embedded in a landscape combined with the excellent representation of the sky
provided by Stellarium. You can walk around, check for (or
demonstrate) possible astronomical alignments of ancient architecture, see
sundials and other shadow casters in action, etc.

\section{Usage}
\label{sec:scenery3d:Usage}


You activate the plugin with the \emph{circular enclosure} button at screen
bottom or by pressing \key{\ctrl+W}. The other button with circular enclosure and
tool icon (or \key{\ctrl+\shift+W}) opens the settings dialog. Once loaded and
displaying, you can walk around pressing \key{\ctrl} plus cursor keys. Change eye
height with \key{\ctrl+Page \arrowkeyup}/\key{\ctrl + Page \arrowkeydown} keys. Adding \key{\shift} key increases speed by 10,
adding \key{\Alt} multiplies by 5 (pressing both keys multiplies by 50!). If you release \key{\ctrl} before
the cursor key, animation will continue. (Press \key{\ctrl}+any cursor key to stop
moving.)\footnote{I (GZ) had to change keyboard handling in the
main program, somewhat breaking the plugin concept.}

Further key bindings exist which can be configured using the Stellarium default
key-binding interface. Some options are also available in the Scenery3d dialog.
For example, coordinate display can be enabled with \key{\ctrl+R}+\key{T}. If your models are georeferenced
in a true geographical coordinate grid, e.g.\ UTM or Gauss-Krueger, you will
especially like this, and this makes the plugin usable for scientific purposes.
Display shows grid name, Easting, Northing, Altitude of ground, and eye height
above ground.

Other features include a virtual ``torchlight'', which can be enabled with \key{\ctrl+R}+\key{L} to give
additional local illumination around the viewer to help to see in the dark.
Interesting points of view can be saved and restored later by the user,
including a description of the view. Scene authors can also distribute
predefined viewpoints in their scene.

The plugin also simulates the shadows of the scene's objects cast by
the Sun, Moon and even Venus (only 1 shadow caster used at a time, you
will never see shadows cast by Venus in moonlight), so you could use
it for examining sundials, or analyze and simulate light-and-shadow
interactions in archaeological structures.

\section{Hardware Requirements \& Performance}
\label{sec:scenery3d:HardwareRequirements}

In order to work with the non-linear projection models in Stellarium,
this plugin uses a trick to create the foreground renderings: it
renders the scene into the six planes of a so-called cubemap, which is
then correctly reprojected onto the sides of a cube, depending on the
current projection settings. Your graphics card must be able to do
this, i.e. it must support the OpenGL extension called
\texttt{EXT\_framebuffer\_object}. Typical modern 3D cards (by NVidia
or ATI/AMD) support this extension. In case your graphics hardware
does not suppport it, the plugin will still work, but you are limited
to perspective projection.

You can influence rendering quality, but also speed, using the plugin's 
GUI, which provides some options such as enabling the use
of shadows, bumpmapping (provides more realistic surface lighting) or 
configuring the sizes of the textures used
for the cubemap or shadowmaps. Larger values there improve the quality,
but require faster hardware and more video memory for smooth results.

Because the ``cubemap trick'' requires quite a large amount of performance (in
essence, the scene has to be rendered 6 times), there are some options available
that try to reduce this burden. The first option is to change the type of the
``cubemap''. The most compatible setting is \emph{6 textures}, which seems to
work best on older integrated Intel GPUs. The recommended default is the second
setting, \emph{Cubemap}, which uses a more modern OpenGL feature and generally
works a bit faster than \emph{6 textures} on more modern graphics cards. Finally,
the \emph{Geometry shader} option tries to render all 6 cube faces at once. This
requires a more recent GPU + drivers (at least OpenGL 3.2 must be supported),
the setting is disabled otherwise. Depending on your hardware and the scene's
complexity, this method may give a speedup or may be slower, you must find this out yourself.

Another option prevents re-rendering of the cubemap if nothing relevant has
changed. You can define the interval (in Stellarium's simulation time) in which
nothing is updated in the GUI. You can still rotate the camera without causing a
re-draw, giving a subjective performance that is close to Stellarium's
performance without Scenery3d. When moving, the cubemap will be updated. You can
enable another option that only causes 1 or 2 sides of the cubemap to be updated
while you move, giving a speedup but causing some parts of the image to be
outdated and discontinuous. The cubemap will be completed again when you stop
moving.

Shadow rendering may also cause quite a performance impact. The \emph{Simple
shadows} option can speed this up a lot, at the cost of shadow quality
especially in larger scenes. Another performance/quality factor is shadow
filtering. The sharpest (and fastest) possible shadows are achieved with
filtering \emph{Off}, but depending on shadowmap resolution and scene size the
shadows may look quite ``blocky''. \emph{Hardware} shadow filtering is usually
very fast, but may not improve appearance a lot. Therefore, there are additional
filter options available, the \emph{High} filter option is relatively expensive.
Finally, the \emph{PCSS} option allows to approximate the increase of solar and lunar shadow
penumbras relative to the distance from their shadow casters, i.e. shadows are
sharp near contact points, and more blurred further away. This again requires
quite a bit of performance, and only works if the shadow filter option is set to
\emph{Low} or \emph{High} (without \emph{Hardware}).

The configuration GUI shows tooltips for most of its settings, which can explain
what a setting does. All settings are saved automatically, and restored when you
reopen Stellarium.


\subsection{Performance notes}
\label{sec:scenery3d:Performance}

On reasonably good hardware 
%2012: (tested on a notebook PC with NVidia M9800 GTS), 
% models up to 100.000 triangles are fluent, up to 250.000 are still ``interactive''.  
% we have 2015, and a much faster implementation!
(tested on a notebook PC with NVidia M580 GTS), models with about
500.000 triangles are fluent with shadows and bumpmaps.  On very small
hardware like single-board computers with native OpenGL ES2, models
may be limited to 64k vertices (points).  If display is too slow,
switch to perspective projection: all other projections require almost
sixfold effort!  You should also prefer the ``lazy'' cubemap mode, 
where the scene is only rendered in specific timesteps or
when movement happens.  


\section{Model Configuration}
\label{sec:scenery3d:ModelConfiguration}

The model format supported in Scenery3d is Wavefront .OBJ, which is
pretty common for 3D models.  You can use several modeling programs to
build your models. Software such as \program{Blender}, \program{Maya}, \program{3D Studio
Max} etc.\ can export OBJ. 

\subsection{Exporting OBJ from Sketchup}
\label{sec:scenery3d:sketchup}

A simple to use and cost-free modeling program is \program{Sketchup}, commonly
used to create the 3D buildings seen in \program{Google Earth}. It can be used
to create georeferenced models.  OBJ is not a native export format for
the standard version of \program{Sketchup}. If you are not willing to
afford \program{Sketchup Pro}, you have to find another way to export a textured
OBJ model.


\begin{table}[t]
  \centering
\begin{tabular}{rl}\toprule
Geometry&Yes\\Lights&Yes\\Clay&No\\Photomatched&Yes\\DefaultUVs&No\\Instanced&No\\\bottomrule
\end{tabular}
\caption{Kerkythea Export Settings}
  \label{fig:scenery3d:KerkytheaExportSettings}
\end{table}

One good exporter is available in the Kerkythea renderer project\footnote{
Available at \url{http://www.kerkythea.net/cms/}}.  You need
\program{SU2KT~3.17}
or better, and \program{KT2OBJ~1.1.0} or better.  Deselect any selection, then
export your model to the Kerkythea XML format with settings shown in
\ref{fig:scenery3d:KerkytheaExportSettings}.
(Or, with selection enabled, make sure settings are No-Yes-Yes-No-Yes-No-No.)  
You do not have to launch Kerkythea unless you want to create nice renderings of
your model. 
Then, use the \program{KT2OBJ} converter to create an OBJ.  You can delete the
XML after the conversion.  Note that some texture coordinates may not be
exported correctly. The setting Photomatched:Yes seems now to have
corrected this issue, esp. with distorted/manu\-ally shifted textures.

Another free OBJ exporter has been made available by 
TIG: \file{OBJexporter.rb}\footnote{Available from
\url{http://forums.sketchucation.com/viewtopic.php?f=323&t=33448}}. 
This is the only OBJ exporter tested so far capable of handling large TIN landscapes
($>450.000$ triangles). 
As of version 2.6 it seems to be the best OBJ exporter available for Sketchup. 

%As of version 1.6 it appears to have valid texture coordinates, but I have
%experienced problems with black faces which need more
%investigation. Maybe you can combine two OBJ files with another 3D
%editor after creating individual OBJs if this is a problem. 

This exporter swaps Y/Z coordinates, but you can add a key to the config file to
correct swapped axes, see below. Other exporters may also provide coordinates in
any order of X, Y, Z -- all those can be properly configured.

Another (almost) working alternative: \file{ObjExporter.rb} by author
Honing.  Here, export with settings \texttt{0xxx00}. This will not create a
\file{TX...} folder but dump all textures in the same directory as the OBJ
and MTL files. Unfortunately, currently some material assignments seem to be
bad.

Yet another exporter, \file{su2objmtl}, does also not provide good texture
coordinates and cannot be recommended at this time.

\subsection{Notes on OBJ file format limitations}
\label{sec:scenery3d:OBJlimitations}

The OBJ format supported is only a subset of the full OBJ format: Only
(optionally textured) triangle meshes are supported, i.e., only lines containing
statements: \parameter{mtllib}, \parameter{usemtl}, \parameter{v}, \parameter{vn}, \parameter{vt}, \parameter{f}
(with three elements only!), \parameter{g}. Negative vertex numbers (i.e., a
specification of relative positions) are not supported.

A further recommendation for correct illumination is that all vertices should
have vertex normals. Sketchup models exported with the Kerkythea or TIG plugins
should have correct normals. If your model does not provide them, default
normals can be reconstructed from the triangle edges, resulting in a faceted
look.

If possible, the model should also be triangulated, but the current loader may
also work with non-triangle geometry. 
%% TODO: GZ: Does that mean Quads, or higher-level polygons? 
The correct use of objects (\parameter{o}) and
groups (\parameter{g}) will improve performance: it is best if you pre-combine all objects
that use the same material into a single one. The loader will try to optimize it
anyways if this is not the case, but can do this only partly (to combine 2
objects with the same material into 1, it requires them to follow directly after
each other in the OBJ). A simple guide to use \program{Blender} for this task follows:

\begin{enumerate}
\item Import from Wavefront (.obj) - you may need to change the forward/up axes for correct orientation, try -Y for forward and Z for up
\item Select an object which has a shared material
\item Press \key{\shift+L} and select 'By Material'
\item Select 'Join' in the left (main) tool window
\item Repeat for other objects that have shared materials
\item Export the .obj, making sure to select the same forward/up axes as in the import, also make sure ``Write Normals'', ``Write Materials'' and ``Include UVs'' are checked
\end{enumerate}

\noindent For transparent objects (with a \parameter{d} or \parameter{Tr} value, alpha testing does
NOT need this), this recommendation does NOT hold: for optimal
results, each separate transparent object should be exported as a
separate ``OBJ object''. This is because they need to be sorted during
rendering to achieve correct transparency. If the objects are combined
already, you can separate them using Blender:

\begin{enumerate}
\item Import .obj (see above)
\item Select the combined transparent object
\item Enter ``Edit'' mode with \key{\tab} and make sure everything is selected (press \key{A} if not)
\item Press \key{P} and select ``By loose parts'', this should separate the object into its unconnected regions
\item Export .obj (see above), also check ``Objects as OBJ Objects''
\end{enumerate}

\begin{table}[tb]
\begin{tabular}{llll}
\toprule
\emph{Parameter}  &\emph{Default} &\emph{Range}    & \emph{Meaning}\\
\midrule
\parameter{Ka}            &set to \parameter{Kd} values & $0\dots1$ each& R/G/B Ambient color\\
\parameter{Kd}            &0.8 0.8 0.8  & $0\dots1$ each & R/G/B Diffuse color\\
\parameter{Ke}            &0.0 0.0 0.0  & $0\dots1$ each & R/G/B Emissive color\\
\parameter{Ks}            &0.0 0.0 0.0  & $0\dots1$ each & R/G/B Specular color\\
\parameter{Ns}            &8.0          & $0\dots\infty$ & shinyness \\
\parameter{d} or \parameter{Tr} &1.0          & $0\dots1$      & opacity \\
\parameter{bAlphatest}    &0            & 0 or 1         & perform alpha test \\
\parameter{bBackface}     &0            & 0 or 1         & render backface \\
\parameter{map\_Kd}       & (none)      & filename       & texture map to be mixed with Ka, Kd \\
\parameter{map\_Ke}       & (none)      & filename       & texture map to be mixed with Ke \\
\parameter{map\_bump}     & (none)      & filename       & normal map for surface roughness\\  
\bottomrule
\end{tabular}
\caption{MTL parameters evaluated}
\label{tab:scenery3d:MTL}
\end{table}

\noindent The MTL file specified by \parameter{mtllib} contains the material parameters. The
minimum that should be specified is either \parameter{map\_Kd} or a \parameter{Kd} line
specifying color values used for the respective faces. But there are other
options in MTL files, and the supported parameters and defaults are listed in
Table~\ref{tab:scenery3d:MTL}.

If no ambient color is specified, the diffuse color values are taken for the
ambient color. An optional emissive term \parameter{Ke} can be added, which is
modulated to only be visible during nighttime. This also requires the
landscape's self-illumination layer to be enabled. It allows to model
self-illuminating objects such as street lights, windows etc. It can optionally
also be modulated by the emissive texture \parameter{map\_Ke}.

If a value for \parameter{Ks} is specified, specularity is evaluated
using the Phong reflection
model\footnote{\url{https://en.wikipedia.org/wiki/Phong_reflection_model}}
with \parameter{Ns} as the exponential shininess constant. Larger
shininess means smaller specular highlights (more metal-like
appearance). Specularity is not modulated by the texture maps.

If a value for \parameter{d} or \parameter{Tr} exists, alpha blending is enabled for this
material. This simulates transparency effects. Transparency can be further
controlled using the alpha channel of the \parameter{map\_Kd} texture.

A simpler and usually more performant way to achieve simple ``cutout''
transparency effects is alpha-testing, by setting \parameter{bAlphatest} to 1. This
simply discards all pixels of the model where the alpha value of the
\parameter{map\_Kd} is below the \parameter{transparency\_threshold} value from
\file{scenery3d.ini}, making ``holes'' in the model. This also produces
better shadows for such objects. If required, alpha testing can be combined with
``real'' blending-based transparency.

Sometimes, exported objects only have a single side (``paper wall''), and are only visible
from one side when looked at in Scenery3d. This is caused by an optimization
called \href{https://en.wikipedia.org/wiki/Back-face_culling}{back-face
culling}, which skips drawing the back sides of objects because they are usually
not visible anyway. If possible, avoid such ``thin'' geometry, this will also
produce better shadows on the object. As a workaround, you can also set
\parameter{bBackface} to 1 to disable back-face culling for this material.

The optional \parameter{map\_bump} enables the use of a tangent-space
normal
maps\footnote{\url{https://en.wikipedia.org/wiki/Normal_mapping}},
which provides a dramatic improvement in surface detail under
illumination.

\subsection{Configuring OBJ for Scenery3d}
\label{sec:scenery3d:Configuring}

The walkaround in your scene can use a ground level (piece of terrain)
on which the observer can walk. The observer eye will always stay ``eye
height'' above ground. Currently, there is no collision detection with
walls implemented, so you can easily walk through walls, or jump on
high towers, if their platform or roof is exported in the ground
layer. If your model has no explicit ground layer, walk will be on the
highest surface of the scenery layer.  If you use the special name
\parameter{NULL} as ground layer, walk will be above \parameter{zero\_ground\_height} level.

Technically, if your model has cavities or doors, you should export
your model twice. Once, just the ground plane, i.e. where you will
walk. Of course, for a temple or other building, this includes its
socket above soil, and any steps, but pillars should not be included.  
This plane is required to compute
eye position above ground. Note that it is not possible to walk in
several floors of a building, or in a multi-plane staircase. You may
have to export several ``ground'' planes and configure several scenery
directories for those rare cases. For optimal performance, the ground 
model should consist of as few triangles as you can tolerate.

The second export includes all visible model parts, and will be used for
rendering. Of course, this requires the ground plane again, but also
all building elements, walls, roofs, etc. 

If you have not done so by yourself, it is recommended to separate
ground and buildings into Sketchup layers 
(or similar concepts in whichever editor you are using) 
in order to easily switch the model to the right state prior to exporting.

Filename recommendations: 


\begin{longtabu}to \textwidth {lX} 
\file{<Temple>.skp}         &Name of a Sketchup Model file. 
                             (The \file{<>} brackets signal ``use your own name here!'')
                             The SKP file is not used by Scenery3d, but you may want
                             to leave it in the folder for later improvements.\\
\file{<Temple>.obj}         &Model in OBJ format. \\
\file{<Temple>\_ground.obj} &Ground layer, if different from Model file. 
\end{longtabu}

\noindent OBJ export may also create folders \file{TX\_<Temple>} and
\file{TX\_<Temple>\_ground}. You can delete the \file{TX\_<Temple>\_ground} folder, 
\file{<Temple>\_ground.obj} is just used to compute vertical height.

%Stellarium uses a directory to store additional data per-user. On Windows, this
%defaults to \verb|C:\Documents and Settings\<username>\Application Data\Stellarium|, 
%but you can use another directory by using the command-line
%argument \parameter{--user-dir <USERDATA>}.  We will refer to this directory. 

Put the
OBJ, MTL and TX directories into a subdirectory of your user directory (see section~\ref{sec:Directories}), e.g.\ 
\file{<USERDATA>/Stellarium/scenery3d/<Temple>}, and add a text file into it 
called \file{scenery3d.ini} (This name is mandatory!) with content described as
follows.

% A Sketchup plugin "Write scenery3d.ini for Stellarium" will write this
% file. Locate the directory where the .obj file(s) reside(s), and store
% scenery3d.ini there. If you have other modelers and models, or if your
% model is not georeferenced in Sketchup, write the file yourself and
% use the following format.
%
% TBD GZ: Release this Sketchup export plugin?

\begin{longtabu}to \textwidth {lX} 
\texttt{[model]}&\\
\texttt{name=<Temple>}&      Unique ID within all models in scenery3d directory.
                             Recommendation: use directory name.\\
\texttt{landscape=<landscapename>}& Name of an available Stellarium landscape.
\end{longtabu}

\noindent This is required if the landscape file includes geographical
coordinates and your model does not: First, the location coordinates
of the \file{landscape.ini} file are used, then location coordinates given here.
The landscape also provides the background image of your scenery. If
you want a zero-height (mathematical) horizon, use the provided
landscape called \file{Zero Horizon}.

\begin{longtabu}to \textwidth {lX} 
\texttt{scenery=<Temple>.obj}         &The complete model, including visible ground.\\
\texttt{ground=<Temple>\_ground.obj}  &Optional: separate ground plane. (NULL for zero altitude.)\\
\texttt{description=<Description>}    &A basic scene description (including HTML tags)
\end{longtabu}

\noindent The \file{scenery3d.ini} may contain a simple scene description, but it is
recommended to use the \emph{localizable} description format: in the scene's
directory (which contains \file{scenery3d.ini}) create files in the format
\file{description.<lang\_code>.utf8} which can contain arbitrary
UTF-8--encoded HTML content. \parameter{<lang\_code>} stands for the ISO~639
language code.

\begin{longtabu}to \textwidth {lX} 
\multicolumn{2}{l}{\texttt{author=<Your Name yourname@yourplace.com>}}\\
\multicolumn{2}{l}{\texttt{copyright=<Copyright Info>}}\\
\texttt{obj\_order=XYZ}      & Use this if you have used an exporter which swaps Y/Z coordinates. 
                               Defaults to \texttt{XYZ}, other options: \texttt{XZY}, \texttt{YZX}, \texttt{YXZ}, \texttt{ZXY}, \texttt{ZYX}\\
\texttt{camNearZ=0.3}        & This defines the distance of the camera near plane, default 0.3.
                               Everything closer than this value to the camera can not be 
                               displayed. Must be larger than zero. It may seem tempting 
                               to set this very small, but this will lead to accuracy issues. 
                               Recommendation is not to go under 0.1\\
\texttt{camFarZ=10000}       & Defines the maximal viewing distance, default 10000.\\
\texttt{shadowDistance=<val>}& The maximal distance shadows are displayed. If left out, the
                               value from \texttt{camFarZ} is used here. If this is set to a smaller
                               value, this may increase the quality of the shadows that are
                               still visible.\\
\texttt{shadowSplitWeight=0..1}& Decimal value for further shadow tweaking. If you require
                                 better shadows up close, try setting this to higher values.
                                 The default is calculated using a heuristic that incorporates
                                 scene size.\\
\texttt{[general]}&\\
\end{longtabu}

\noindent The general section defines some further import/rendering options.
% GZ TBD: Why do we need ground normals? 
%
\begin{longtabu}to \textwidth {lX} 
\texttt{transparency\_threshold=0.5}  &Defines the alpha threshold for alpha-testing, as described in section~\ref{sec:scenery3d:OBJlimitations}. Default \texttt{0.5}\\
\texttt{scenery\_generate\_normals=0} &Boolean, if true normals are recalculated by the plugin, instead of imported. Default \texttt{false}\\
\texttt{ground\_generate\_normals=0}  &Boolean, same as above, for ground model. Default \texttt{false}.\\
\texttt{[location]}                   &
\end{longtabu}

\noindent Optional section to specify geographic longitude $\lambda$, latitude $\varphi$,
and altitude. Required if \verb|coord/convergence_angle=from_grid|, else
location is inherited from landscape.
\begin{longtabu}to \textwidth {lX} 
\multicolumn{2}{l}{\texttt{planet = Earth}}\\
\texttt{latitude = +48d31'30.4"}      & Required if \texttt{coord/convergence\_angle=from\_grid}\\
\texttt{longitude = +16d12'25.5"}     & "--"\\
\texttt{altitude =from\_model|<int>}   &  altitude (for astronomical computations) can be computed from the model:  if
                                           \texttt{from\_model}, it is computed as $(z_{min}+z_{max})/2+\mathtt{orig\_H}$, 
                                            i.e.\ from the model bounding box centre height.\\
\multicolumn{2}{l}{\texttt{display\_fog = 0}}\\
\multicolumn{2}{l}{\texttt{atmospheric\_extinction\_coefficient = 0.2}}\\
\multicolumn{2}{l}{\texttt{atmospheric\_temperature = 10.0}}\\
\multicolumn{2}{l}{\texttt{atmospheric\_pressure = -1}}\\
\multicolumn{2}{l}{\texttt{light\_pollution = 1}}\\
\multicolumn{2}{l}{\texttt{[coord]}}\\
\end{longtabu}

\noindent Entries in the \verb|[coord]| section are again optional,
default to zero when not specified, but are required if you want to
display meaningful eye coordinates in your survey (world) coordinate
system, like UTM or Gauss-Kr\"uger.

\begin{longtabu}to \textwidth {lX} 
\texttt{grid\_name=<string> }& Name of grid coordinates, e.g. \texttt{"UTM 33 U (WGS 84)"},
                               \texttt{"Gauss-Kr\"uger M34"} or \texttt{"Relative to <Center>"} This name is
                              only displayed, there is no evaluation of its contents.\\
\texttt{orig\_E=<double> | (Easting)}  &East-West-distance to zone central meridian\\
\texttt{orig\_N=<double> | (Northing)} &North distance from Equator\\
\texttt{orig\_H=<double> | (Height)}   &Altitude above Mean Sea Level of model origin\\
\end{longtabu}

\noindent These entries describe the offset, in metres, of the model coordinates relative
to coordinates in a geographic grid, like Gauss-Kr\"uger. If you have your model
vertices specified in grid coordinates, do not specify \verb|orig_...| data, but
please definitely add \verb|start_...| data, below.

Note that using grid coordinates without offset for the vertices is
usually a bad idea for real-world applications like surveyed sites in
UTM coordinates. Coordinate values are often very large numbers
(ranging into millions of meters from equator and many thousands from
the zone meridian). If you want to assign millimetre values to model
vertices, you will hit numerical problems with the usual
single-precision floating point arithmetic. Therefore we can specify
this offset which is only necessary for coordinate display.

Typically, digital elevation models and building structures built on those are
survey-grid aligned, so true geographical north will not coincide with grid
north, the difference is known as \indexterm{meridian convergence}\footnote{%
  \url{http://en.wikipedia.org/wiki/Transverse_Mercator_projection}}.
\begin{equation}
\gamma(\lambda, \varphi)=\arctan(\tan(\lambda-\lambda_0)\sin\varphi)
\end{equation}
This amount can be given in \verb|convergence_angle| (degrees), so
that your model will be rotated clockwise by this amount around the
vertical axis to be aligned with True North\footnote{Note that
  Sketchup's \emph{georeferencing dictionary} provides a NorthAngle
  entry, which is $360-\mathtt{convergence\_angle}$.}.

\begin{verbatim}
convergence_angle=from_grid|<double> 
grid_meridian=<double>|+<int>d<int>'<float>"      
\end{verbatim}


\noindent \texttt{grid\_meridian} is the central meridian $\lambda_0$ of grid
zone, e.g.\ for UTM or Gauss-Kr\"uger, is only required to compute
convergence angle if \texttt{convergence\_angle=from\_grid}



\begin{verbatim}
zero_ground_height=<double> 
\end{verbatim}
height of terrain outside \file{<Temple>\_ground.OBJ}, or if \verb|ground=NULL|.
Allows smooth approach from outside.  This value is relative to the model
origin, or typically close to zero, i.e.,  use a Z value in model coordinates,
not world coordinates! (If you want the terrain height surrounding your model to
be \texttt{orig\_H}, use 0, not the correct mean height above sea level!)  Defaults
to minimum of height of ground level (or model, resp.) bounding box.

\begin{longtabu}to \textwidth {lX} 
\multicolumn{2}{l}{\texttt{start\_E=<double>}}\\ 
\multicolumn{2}{l}{\texttt{start\_N=<double>}} \\
\texttt{start\_H=<double>}     &only meaningful if \texttt{ground==NULL}, else \texttt{H} is derived from ground \\
\texttt{start\_Eye=<double>}   &default: \texttt{1.65m} \\
\multicolumn{2}{l}{\texttt{start\_az\_alt\_fov=<az\_deg>,<alt\_deg>,<fov\_deg>}}\\
\                                &initial view direction and field of view.\\
\end{longtabu}

\noindent \texttt{start\_...} defines the view position to be set after loading the scenery.
Defaults to center of model boundingbox.

It is advisable to use the grid coordinates of the location of the panoramic
photo (landscape) as \texttt{start\_...} coordinates, or the correct coordinates
and some carefully selected \texttt{start\_az\_alt\_fov} in case of certain view
corridors (temple axes, \ldots).

\subsection{Predefined views}
You can also distribute some predefined views with your model in a
\file{viewpoints.ini} file. See the provided ``Sterngarten'' scene for an
example. These entries are not editable by the user through the interface. The
user can always save his own views, they will be saved into the file
\file{userviews.ini} in the user's Stellarium user directory, and are editable.

\begin{configfile}[\scriptsize]
[StoredViews]
size=<int>              Defines how many entries are in this file. 
                        Prefix each entry with its index!
1/label=<string>        The name of this entry
1/description=<string>  A description of this entry (can include HTML)
1/position=<x,y,z,h>    The x,y,z grid coordinates 
                        (like orig_* or start_* in scenery3d.ini) 
                        + the current eye height h
1/view_fov=<az_deg,alt_deg,fov_deg>  The view direction + FOV
                                     (like start_az_alt_fov in scenery3d.ini)
; an example for a second entry (note the 2 at the beginning of each line!)
2/label       = Signs
2/description = Two signs that describe the Sterngarten
2/position    = 593155.2421,5333348.6304,325.7295809038,0.8805
2/view_fov    = 84.315399,-8.187078,83.000000
\end{configfile}

\subsection{Concatenating OBJ files}
\label{sec:scenery3d:concatenating}

Some automated workflows may involve tiled landscape areas, e.g. to
overcome texture limitations or triangle count limits in simpler tools
like \program{Sketchup}. In this case you can create separate meshes in the same
coordinate system, but you need to concatenate them. in \program{Blender}, import
the OBJ files \menu{File>Import>Wavefront .obj}, select them and press
\key{\ctrl+J} to join them, then export to a single OBJ
\menu{File>Export>Wavefront .obj}%
\footnote{http://blender.stackexchange.com/questions/3352/merging-multiple-obj-files}.

Verify the new model loads correctly, e.g.\ in Meshlab!

\subsection{Working with non-georeferenced OBJ files}
\label{sec:scenery3d:NonGeoreferenced}


There exists modeling software which produces nice models, but without
concept of georeference. One spectacular example is \program{AutoDesk PhotoFly},
a cloud application which delivers 3D models from a bunch of photos
uploaded via its program interface. This ``technological preview'' is
in version 2 and free of cost as of mid-2011.

The problem with these models is that you cannot assign surveyed
coordinates to points in the model, so either you can georeference the
models in other applications, or you must find the correct
transformation matrix.  Importing the OBJ in \program{Sketchup} may take a long
time for detailed photo-generated models, and the texturing may
suffer, so you can cut the model down to the minimum necessary e.g.\ in
Meshlab, and import just a stub required to georeference the model in
\program{Sketchup}. 

Now, how would you find the proper orientation? The easiest chance
would be with a structure visible in the photo layer of Google
Earth. So, start a new model and immediately \menu{add location} from the
Google Earth interface. Then you can import the OBJ with TIG's importer
plugin.  If the imported model looks perfect, you may just place the
model into the Sketchup landscape and export a complete landscape just
like above. If not, or if you had to cut/simplify the OBJ to be able
to import it, you can rotate/scale the OBJ (it must be grouped!). If
you see a shadow in the photos, you may want to set the date/time of
the original photos in the scene and verify that the shadows created by
\program{Sketchup} illuminating the model match those in the model's photo
texture. When you are satisfied with placement/orientation, you create
a \file{scenery3d.ini} like above with the command
\menu{Plugins>ASTROSIM/Stellarium scenery3d helpers>Create scenery3d.ini}.

Then, you select the OBJ group, open \menu{Windows>Ruby Console} and call
\menu{Plugins>ASTROSIM/Stellarium scenery3d helpers>Export transformation
of selected group}.

On the Ruby console, you will find a line of numbers (the $4\times4$
transformation matrix) which you copy/paste (all in one line!) into the
\file{[model]} section in \file{scenery3d.ini}.
\begin{verbatim}
obj2grid_trafo=<a11>,<a12>,<a13>,<a14>,<a21>,<a22>,<a23>,<a24>,
               <a31>,<a32>,<a33>,<a34>,<a41>,<a42>,<a43>,<a44>
\end{verbatim}
You edit the \file{scenery3d.ini} to use your full (unmodified)
\program{PhotoFly} model and, if you don't have a panorama, take \parameter{Zero Horizon}
landscape as (no-)background. It depends on the model if you want to
be able to step on it, or to declare \texttt{ground=NULL} for a
constant-height ground. Run Stellarum once and adjust the
\texttt{start\_N}, \texttt{start\_E} and \texttt{zero\_ground\_height}.

\subsubsection{Rotating OBJs with recognized survey points}
\label{sec:scenery3d:RotatingOBJ}

If you have survey points measured in a survey grid plus a photomodel
with those points visible, you can use Meshlab to find the model
vertex coordinates in the photo model, and some other program like
\program{CoordTrans} in the \program{JavaGraticule3D} suite to find either the matrix
values to enter in \file{scenery3d.ini} or even rotate the OBJ
points. However, this involves more math than can be described here;
if you came that far, you likely know the required steps.  Here it
really helps if you know how to operate automatic text processors like
\program{AWK}.

\section*{Authors and Acknowledgements}
\label{sec:scenery3d:Acknowledgments}


Scenery3d was conceived by Georg Zotti for the
ASTROSIM\footnote{\url{http://astrosim.univie.ac.at}} project. A first
prototype was implemented originally in 2010/2011 by Simon Parzer and
Peter Neubauer as student work supervised by Michael Wimmer (TU Wien).
Models for accuracy tests (Sterngarten, Testscene), and later
improvements in integration, user interaction, \file{.ini} option handling,
OBJ/MTL loader bugfixes and georeference testing by Georg Zotti.

Andrei Borza in 2011/12 further improved
rendering quality (shadow mapping, normal mapping) and speed. 

In 2014/15,
Florian Schaukowitsch adapted the code to work with Qt 5 and the 
Stellarium 0.13 codebase, replaced the renderer with a more efficient, fully
shader-based system, implemented various performance, quality and usability
enhancements, and did some code cleanup. Both Andrei and Florian were again
supervised by Michael Wimmer.

This work has been originally created during the ASTROSIM project supported 2008-2012 by
the Austrian Science Fund (FWF) under grant number P~21208-G19.

%\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "guide"
%%% End: 

